### 2019-07-31 Manuscript.Rmd
Imported .Rmd file

---
title: 
author: Tina & Marc
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook:
    fig_caption: yes
    keep_tex: yes
    highlight: espresso
    number_sections: no
  html_document:
    df_print: paged
  word_document:
    reference_docx: Paper1_reply_v13_Tina_Marc.docx
urlcolor: blue
fontsize: 10pt
geometry: margin=1in
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{float}
- \usepackage{color}
linestretch: 1
bibliography: /Users/admin/Desktop/nwork/simulation/TM.bib
csl: /Users/admin/Desktop/nwork/simulation/styles-master/styles-master/plos.csl 
---

<style>
body {
text-align: justify}
</style>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE,message=FALSE, warning=FALSE, dev="png",dpi = 400,cache = TRUE)
```

```{r Initial settings, echo=F, message=F, warning=F}
  # The following libraries should be installed from github with the specified 
	# devtools command. One needs to install devtools first.
	library(treeio)#devtools::install_github("GuangchuangYu/treeio")
  library(ggtree)#devtools::install_github("GuangchuangYu/ggtree")
  library(easyGgplot2) #install_github("kassambara/easyGgplot2")

	# The following libraries can be installed from CRAN by using  
  #install.packages("packagename")
  library(geiger)
  library(ape)
  library(phytools)
  library(stringr)
  library(ggrepel)
  library(dplyr)
  library(ggplot2)
  library(gridExtra)
  library(parallel)
  library(digest)
  library(magrittr)
  library(gtools)
  library(tidyverse)
  library(caper)
  library(foreach)
  library(doParallel)
  library(cowplot)
  library(png)
  library(grid)

  ## Setting the working directory 
  setwd("/Users/admin/Desktop/nwork/simulation/")
  set.seed(23456)

  ## Renamed original function file provided by Dunn et al. (doi:10.1073/pnas.1707515115)
  source("functions_Dunn.R")

  ## Additional functions written for this study
  source( "/Users/admin/Desktop/nwork/Empirical/protein_trees.nhx/function_TM_new.R" )

  ## Setting system computational parameter
	cores = detectCores()

```

\vspace{0.5cm}

# **Phylogenetic comparative methods are problematic when applied to gene trees with speciation and duplication nodes: correcting for biases in testing the ortholog conjecture**

\vspace{0.5cm}
Tina Begum^1,2^, Marc Robinson-Rechavi^1,2^*

\vspace{0.5cm}

^1^Department of Ecology and Evolution, University of Lausanne, Lausanne, Switzerland<br />      
^2^SIB Swiss Institute of Bioinformatics, Lausanne, Switzerland <br />

\vspace{0.5cm}

\*Corresponding author  
E-mail: marc.robinson-rechavi@unil.ch

### Abstract

Most comparative studies of functional genomics have used pairwise comparisons. Yet it has been shown that this can provide biased results, since genes, like species, are phylogenetically related. Phylogenetic comparative methods should allow to correct for this, but they depend on strong assumptions, including unbiased tree estimates relative to the hypothesis being tested. An ongoing trend in comparative genomic studies is to adopt phylogenetic comparative method to answer a wide range of biological questions including evolutionary hypotheses testing. Notably among them is the recently controversial ???Ortholog Conjecture??? that assumes the functional evolution is faster in paralogs than orthologs. Using pairwise comparisons of tissue specificity index ($\tau$), earlier we provided support for the ortholog conjecture. In contrast, a recent publication suggested that the ortholog conjecture is not supported by gene expression tissue-specificity using phylogenetic independent contrasts. We find that the gene trees used suffer from important biases, due to the inclusion of trees with no duplication nodes, to the relative age of speciations and duplications, to systematic differences in branch lengths, and to non-Brownian motion of tissue-specificity on many trees. We find some support for the ortholog conjecture, but especially that incorrect implementation of phylogenetic method in empirical gene with duplications can be problematic.

### Author Summary

There is a lot of interest in understanding the evolution of gene function. Comparing functional genomics results offers a way to do this. In most cases, it is done by comparing pairs of genes, and notably pairs of orthologs (homologs derived by speciation) and pairs of paralogs (homologs derived by duplication). A drawback of such pairwise comparisons is that they neglect the evolutionary history of the genes, and thus violate the statistical presumption of independence of observations. Phylogenetic comparative methods have been suggested to correct for this. Recent results indicate that whereas pairwise comparisons support stronger functional divergence of paralogs than of orthologs, phylogenetic analyses would not. Re-analyses of these data show that gene trees are biased in such a way as to cause biases in phylogenetic methods when there are gene duplication. This can led to erroneous results, and have serious consequences in biological data interpretation.

\pagebreak

### Introduction

The "Ortholog Conjecture", a cornerstone of phylogenomics, has become a topic of debate in recent years [@RN771;@RN774;@RN790;@RN793;@RN794;@RN762;@RN760;@RN770;@RN792]. Dealing with the roles of gene duplications in functional evolution, the ortholog conjecture is routinely used by both experimental and computational biologists in predicting, or understanding gene function. According to this model, orthologs (i.e. homologous genes which diverged by a speciation event) retain equivalent or very similar functions, whereas paralogs (i.e. homologous genes which diverged by a duplication event) share less similar functions [@RN771]. This is linked to the hypothesis that paralogs evolve more rapidly. This hypothesis was challenged by results suggesting that paralogs would be functionally more similar than orthologs [@RN774]. Such findings not only raised questions on the evolutionary role of gene duplication but also questioned the reliability of using orthologs to annotate unknown gene functions in different species [@RN844]. Several more recent studies [@RN770;@RN790;@RN793;@RN794;@RN762] found support for ortholog conjecture, mostly based on comparisons of gene expression data.

\vspace{0.05cm}
While all previous studies of the ortholog conjecture had used pairwise comparisons of orthologs and paralogs, a recent article suggested that this was flawed, and that phylogenetic comparative methods should be used [@RN760]. In the field of comparative biology, most of the studies still rely on the pairwise method including the testing of the ortholog conjecture. Ignorance of phylogenetic structure has been reported to underestimate the fundamental assumption of independent observations in statistics [@RN773;@RN807]. A solution is to use phylogeny-based methods to investigate such evolutionary patterns [@RN782;@RN807;@RN779;@RN780;@RN773;@RN850;@RN849]. There are three main such phylogenetic methods: Phylogenetic Independent Contrast (PIC) [@RN773], Phylogenetic Generalized Least-Square (PGLS) [@RN850], and Monte Carlo computer simulation [@RN849]. PIC is widely adopted for its relative simplicity, and its applicability to a wide range of statistical procedures [@RN781;@RN760]. The performance of PIC relies on three basic assumptions: a correct tree topology; accurate branch lengths; and trait evolution following a Brownian model (where trait variance accrues as a linear function of time)  [@RN781;@RN782;@RN783;@RN807;@RN780;@RN779;@RN816]. If any of these assumptions is incorrect, this can lead to incorrect interpretation of results. This is probably why the application of such phylogenetic methods is still limited, and debated even after being introduced about three decades ago [@RN781;@RN782;@RN783;@RN807;@RN780]. Dunn et al. [@RN760] took an innovative approach in applying PIC to compare the divergence rates between two different events ("speciation" and "duplication") to test the ortholog conjecture. However, such an application might be problematic since the time of occurrence of gene duplication, one of the two types of events compared, is unknowable by external information (e.g. no fossil evidence). Therefore, further study is required to understand why Dunn et al. [@RN760] obtained results which are inconsistent with most studies by using the phylogenetic method. It is possible that all the conclusions drawn by previous studies on gene duplication are incorrect due to overlooking phylogenetic tree structure. If so, it should be well supported.

\vspace{0.05cm}
We re-examined the data of Dunn et al., after reproducing it using the resources and scripts provided by the authors [@RN760]. Our reanalysis highlights potential problems associated with phylogenetic independent contrasts when applied to the impact of gene duplication. Finally, with proper controls, the phylogenetic method supports the ortholog conjecture.

### Results

```{r load_Dunn_data, echo=F, cache=T, message=F, warning=F}
  ## Loaded the analysis results generated by using files and scripts ("manuscript_kernel.R") of Dunn et al.
  ## Files and scripts are used from the link  https://github.com/caseywdunn/comparative_expression_2017
  ## We provided the following link https://doi.org/10.5281/zenodo.3354285 to make available the reproduced results of Dunn et al. to save time
	load("/Users/admin/Desktop/nwork/simulation/manuscript_dunn.RData")

  ## Obtaining calibrated trees with no duplication event
  gene_trees_with_duplication1<-mclapply(gene_trees_calibrated,tree_with_duplication, mc.cores = cores)
  gene_trees_with_duplication<-gene_trees_with_duplication1[!is.na(gene_trees_with_duplication1)] 
	trees_with_no_duplication <- length(gene_trees_calibrated) - length(gene_trees_with_duplication)

```

__Issues with naive use of Phylogenetic Independent Contrasts (PICs)__

```{r reanalyses_Dunn_data,echo=F, message=F, warning=F}
  ## Dunn et al. replaced the actual Tau values with simulated Tau values (for null and OC simulation) on the calibrated trees 
  ## and returned as treeio::treedata objects, and we used those data for our reanalyses
  ## Creating dataframes of null and OC (Ortholog Conjecture) simulation data using columns of events and PICs  
  null_data.dunn <- NULL
  OC_data.dunn <- NULL
  null_data.dunn <- na.omit(nodes_sim_null_contrast[c(7,21)])
  null_data.dunn$label <- "Null simulation"
  OC_data.dunn <- na.omit(nodes_sim_oc_contrast[c(7,21)])
  OC_data.dunn$label <- "OC simulation"

  ## Retrieving speciation and duplication PICs of null and OC simulation data of Dunn et al.
  speciation_null <- null_data.dunn$pic[which(null_data.dunn$Event=="Speciation")]
  duplication_null <- null_data.dunn$pic[which(null_data.dunn$Event=="Duplication")]
  speciation_oc <- OC_data.dunn$pic[which(OC_data.dunn$Event=="Speciation")]
  duplication_oc <- OC_data.dunn$pic[which(OC_data.dunn$Event=="Duplication")]

  ## Performing Wilcoxon tests on null simulation data 
  wilcox_one_tailed1 <- wilcox.test(duplication_null,speciation_null,alternative = "greater")$p.value
  wilcox_one_tailed2 <- wilcox.test(speciation_null,duplication_null,alternative = "greater")$p.value
  wilcox_two_tailed_null <- two_tailed_wilcox(speciation_null,duplication_null)

  ## Performing two sided Wilcoxon test on OC simulation data 
  wilcox_two_tailed_oc <- two_tailed_wilcox(speciation_oc,duplication_oc)

```

To understand their results, we first reanalyzed the data of Dunn et al. [@RN760]. We were able to reproduce all the results published in their article by running the code, which they clearly supplied. Dunn et al. reported a non-significant result for the PIC under the null simulation, using a Wilcoxon one-tailed rank test to check if the contrasts of duplication events are higher than the contrasts of speciation events (*P* = $`r format(wilcox_one_tailed1, digits= 3, scientific = FALSE)`$). Surprisingly, the PIC rejects the null hypothesis on the null simulations with a Wilcoxon two-tailed rank test (Fig 1A), with significant support for higher contrasts after speciation than duplication. This was robust to repeating the simulations with different random seed number (S1 Fig). This indicates that neither of the methods, PIC or pairwise, worked properly for these calibrated trees, since both rejects the null when simulations are performed under null.

PIC is standardized by the expected variance of daughter branch lengths for each node of event [@RN807;@RN782;@RN779;@RN780]. Therefore, accurate branch length estimation in absolute time (e.g. in Million Years -- My) is needed [@RN781;@RN782;@RN807;@RN779;@RN780]. Since the accuracy and performance of the PIC method largely depend on proper branch length calibration [@RN779;@RN768;@RN782] we investigated possible biases created during calibration of gene trees, especially concerning duplication times for which we have no external reference (e.g. no fossils) (S2 Fig).

```{r phylogenetic_analyses_trees_with_strong_signals, eval=T, echo=F, message=F, warning=F}

  ## Identifying trees with strong phylogenetic signals using the criteria (Blomberg's K > 0.551) as specificed by Dunn et al. in their paper 
  ## We make a vector to collect calibrated tree index with Blomberg's K > 0.551
  tree_index <- as.numeric(row.names(tree_summary)[(!(is.na(tree_summary$K)) & (tree_summary$K > 0.551))])
  trees_new <- gene_trees_calibrated[c(tree_index)] ## List of trees with strong phylogenetic signals
  subset.trees <- length(trees_new)

  ## Counting trees with low phylogenetic signals    
  count_trees <- (length(gene_trees_calibrated)-length(trees_new))/length(gene_trees_calibrated)
  count_trees <- paste(round(count_trees*100, 2), "%", sep="")

  ## Collecting null simulation trees associated with strong phylogenetic signals  
	trees_new_sim_null <- gene_trees_sim_null[c(tree_index)]

	## Making a dataframe by combining data slots of all these trees
	new_sim_null_contrast <- bind_rows(lapply(trees_new_sim_null,summary_function))

	## Collecting OC simulation trees associated with strong phylogenetic signals to make a dataframe with combined data slots for all trees
	trees_new_sim_oc <- gene_trees_sim_oc[c(tree_index)]
	new_sim_oc_contrast <- bind_rows(lapply(trees_new_sim_oc,summary_function))

	## Extracting the columns of event and PIC data from null and OC simulation result of trees with strong phylogenetic signals
  null_data.new <- na.omit(new_sim_null_contrast[c(6,20)])
  null_data.new$label <- "Null simulation"
  OC_data.new <- na.omit(new_sim_oc_contrast[c(6,20)])
  OC_data.new$label <- "OC simulation"

  ## Obtaining speciation and duplication PICs from null and OC simulation of trees with strong phylogenetic signals
  speciation_null_new <- null_data.new$pic[which(null_data.new$Event=="Speciation")]
  duplication_null_new <- null_data.new$pic[which(null_data.new$Event=="Duplication")]
  speciation_oc_new <- OC_data.new$pic[which(OC_data.new$Event=="Speciation")]
  duplication_oc_new <- OC_data.new$pic[which(OC_data.new$Event=="Duplication")]

  ## Performing Wilcoxon tests on null and OC simulation data of trees with strong phylogenetic signals
  wilcox_1_tailed1_new <- one_tailed_wilcox(duplication_null_new,speciation_null_new)
  wilcox_1_tailed2_new <- one_tailed_wilcox(speciation_null_new,duplication_null_new)
  wilcox_2_tailed_null_new <- two_tailed_wilcox(speciation_null_new,duplication_null_new)
  wilcox_2_tailed_oc_new <- two_tailed_wilcox(speciation_oc_new,duplication_oc_new)

```

```{r phylogenetic_signal,eval=T, echo=F, message=F, warning=F}

  ## Identifying time calibrated trees with negative branch lengths
  trees_negative <- negative_edgelength(gene_trees_calibrated)

  ## List of calibrated trees without negative branch lengths
  gene_trees_pic1 <- gene_trees_pic[-trees_negative]

  ## Computing phylogenetic signal for list of trees without negative branch lengths 
  ## Trees for which phylogenetic signal estimation failed, we assigned a random value 1000 to distinguish it from actual estimated values
  signal_tree <- bind_rows(lapply(gene_trees_pic1,Phylogenetic_signal)) 
  signal_tree$tree_number <- as.numeric(rownames(signal_tree))

  ## Dataframe with Blomberg's K
  signal_tree_new <- signal_tree[which(signal_tree$K_effecient<1000),]
  phylosig_blomberg <- signal_tree_new[which(signal_tree_new$Pvalue_Blomberg<0.05),] 
  phylosig_blomberg <- phylosig_blomberg[c(1:3,6)]

  ## Dataframe with Pagel's lamba
  signal_tree_new2 <- signal_tree[which(signal_tree$lambda_estimate<1000),] 
  phylosig_pagel <- signal_tree_new2[which(signal_tree_new2$Pvalue_pagel<0.05),] 
  phylosig_pagel <- phylosig_pagel[c(1,4:6)]

```

The purpose of applying a phylogenetic method is to generate independent statistical data points for traits, which are otherwise not independent enough because of their phylogenetic relatedness. Such statistical non-independence among species trait values can be measured by "phylogenetic signal" [@RN819; @RN768; @RN800; @RN820;@RN826]. Blomberg's K and Pagel's $\lambda$ are widely used methods to capture such phylogenetic signal [@RN819;@RN768]. Dunn et al. [@RN760] used Blomberg's K in their study. In case of Blomberg's K the value ranges from 0 to $\infty$, where a value of 0 indicates no phylogenetic signal, a value of 1 indicates trait evolution according to a Brownian model (BM), values between 0 and 1 indicate that distant relatives are more similar than expected under BM, and values larger than 1 indicate that close relatives are more similar than expected under BM [@RN819; @RN768; @RN800; @RN820;@RN826]. With a cutoff of K > 0.551, Dunn et al. [@RN760] obtained only `r subset.trees` out of `r length(gene_trees_calibrated)` calibrated trees with strong phylogenetic signal. Using a cut-off of *P* < 0.05 together with K > 0.551 leads to 1135 trees, which produce similar results (S3 Fig) to that set of `r subset.trees` trees. We continued analyses with the `r subset.trees` trees of Dunn et al. [@RN760] for consistency. In any case, it is notable that trait ($\tau$) values are independent of phylogeny for the majority of the gene trees. For these `r (length(gene_trees_calibrated)-subset.trees)` trees, implementation of PIC to produce statistically independent data points might not be necessary. Moreover, it can be misleading if the contrasts are not checked for adequate standardization as per BM after applying PIC (discussed later). The phylogenetic method still rejects the null hypothesis under null simulations for those `r subset.trees` trees (Fig 1B), showing that the problem is not simply due to low phylogenetic signal.

```{r Fig1, dpi=400,fig.width=8, fig.height=3.5, fig.cap="**Fig 1**: Reanalyses of phylogenetic simulation data of  Dunn et al. [@RN760]. *P* values are from Wilcoxon two-tailed tests. In null simulations, there should be no difference in contrasts between events. In OC (Ortholog Conjecture) simulations, contrasts are expected to be higher for duplication than for speciation. (A) An excess of contrasts for speciation than duplication reject the null hypothesis under null simulation scenario for all empirical time calibrated gene trees. (B) Results are similar with a subset of trees with strong phylogenetic signal for $\\tau$."}

  ## To make the plot, we made a single dataframe of both simulation data of all calibrated gene trees
  simulation.df <- rbind(null_data.dunn, OC_data.dunn)
  all_trees <- paste0("A. Plots on ",length(gene_trees_calibrated), " trees", sep="")

  ## Dataframe is built based on 2082 trees with strong phylogenetic signals
  simulation.new <- rbind(null_data.new, OC_data.new)
  subset_trees <- paste0("B. Plots on ",subset.trees, " trees", sep="")

  ## Generating plot on simulation data for all 8520 calibrated trees
  dodge <- position_dodge(width = 0.51)

  plotA <- ggplot(simulation.df,aes(x=label, y=pic, fill=Event)) + 
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    xlab(NULL) +
    ylab(expression(bold("Phylogenetic Independent Contrasts (PICs)"))) +
    coord_cartesian(ylim=c(0, 0.05)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9)) +
    theme(legend.position="none")+
    theme(axis.text.y = element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold")) +
    ggtitle(paste0(all_trees))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x=1.04, y=0.035, label=wilcox_two_tailed_null, fontface=4) +
    annotate("text", x=1.98, y=0.035, label=wilcox_two_tailed_oc, fontface=4)

  ## Generating plot on simulation data for all 2082 calibrated trees
  plotB <- ggplot(simulation.new,aes(x=label, y=pic, fill=Event)) + 
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    xlab(NULL) +
    ylab(NULL) +
    coord_cartesian(ylim=c(0, 0.05)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9),legend.text=element_text(size=8)) +
    theme(axis.text.y= element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold"))+
    ggtitle(paste0(subset_trees))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x = 1.04, y = 0.035, label= wilcox_2_tailed_null_new, fontface=4) +
    annotate("text", x = 1.98, y = 0.035, label= wilcox_2_tailed_oc_new, fontface=4)

  grid.arrange(plotA,plotB,ncol = 2)

```

```{r Deatiled_analyses_on_empirical_data, echo=F, message=F, warning=F}

  ## Obtaining calibration time points for speciation and duplication events for empirical data
  speciation.empirical <- unique(round(as.integer(nodes_contrast$node_age[which(nodes_contrast$Event == "Speciation")]), 0))
  duplication.empirical <- unique(round(as.integer(nodes_contrast$node_age[which(nodes_contrast$Event == "Duplication")]), 0))

  ## Expected variance of events
  ## We used a cutoff of pic < 0.5 to get rid of branch length underestimation bias as used by Dunn et al.
  Exp_var.speciation <- nodes_contrast$var_exp[nodes_contrast$Event=="Speciation" & nodes_contrast$pic < 0.5] 
  Exp_var.duplication <- nodes_contrast$var_exp[nodes_contrast$Event=="Duplication"  & nodes_contrast$pic < 0.5] 
  Exp_var.duplication.precede.speciation <- nodes_contrast$var_exp[nodes_contrast$Event=="Duplication" & nodes_contrast$node_age > 296 & nodes_contrast$pic < 0.5] # Expected variance of duplication events preceding oldest speciation event of the trees

  ## Performing test on expected variance
  Exp_var.pvalue <- wilcox.test(Exp_var.duplication.precede.speciation,Exp_var.speciation,alternative="two.sided")$p.value
  if(Exp_var.pvalue == 0) Exp_var.pvalue1 <- 2.2e-16
  if(Exp_var.pvalue != 0) Exp_var.pvalue1 <- Exp_var.pvalue

  ## Exptracting data of trees with strong phylogenetic signals for expected variance analyses 
	trees_new_pic <- gene_trees_pic[c(tree_index)]
	trees_new_contrast <-bind_rows(lapply(trees_new_pic,summary_function))
	trees_new_contrast <-trees_new_contrast[which(!is.na(trees_new_contrast$var_exp)),]
	Exp_var.speciation_subset <- na.omit(trees_new_contrast$var_exp[trees_new_contrast$Event=="Speciation" & trees_new_contrast$pic < 0.5])## all speciation events
  Exp_var.duplication_subset <- na.omit(trees_new_contrast$var_exp[trees_new_contrast$Event=="Duplication"  & trees_new_contrast$pic < 0.5]) ## all duplication events

	 ## Making a dataframe of expected variance for trees with strong phylogenetic signals
  check_df <- data.frame(exp_var=NA,Event=NA)
  check_df <- rbind(check_df, data.frame(exp_var=Exp_var.speciation_subset,Event="Speciation"))
  check_df <- rbind(check_df, data.frame(exp_var=Exp_var.duplication_subset,Event="Duplication"))
  check_df <- check_df[-1,]
  check_df$Event <- factor(check_df$Event, levels=c("Speciation", "Duplication"))

  ## Making a dataframe of expected variance distribution of all 8520 calibrated trees 
  check_df1 <- data.frame(exp_var=NA,Event=NA)
  check_df1 <- rbind(check_df1, data.frame(exp_var=Exp_var.speciation,Event="Speciation"))
  check_df1 <- rbind(check_df1, data.frame(exp_var=Exp_var.duplication,Event="Duplication"))
  check_df1 <- check_df1[-1,]
  check_df1$Event <- factor(check_df1$Event, levels=c("Speciation", "Duplication"))

```

Dunn et al. [@RN760] used `r nrow(calibration_times_focal)` speciation time points to calibrate gene trees. The oldest speciation calibration was at `r max(calibration_times_focal$age)` My.  All other calibrated nodes were duplication nodes leading to obtain ~`r length(duplication.empirical)` unique duplication time points for the `r length(gene_trees_calibrated)` calibrated trees [@RN760]. Out of these, `r length(duplication.empirical[duplication.empirical > max(speciation.empirical)])` time points preceded the oldest speciation node age. Surprisingly, the calibrated node age of the oldest duplication event was found to be `r as.integer(max(duplication.empirical))` My, that is, 2600 times older than the Earth. This is indicative of the difficulty of estimating the age of ancient duplications by phylogenetic methods. Those `r length(duplication.empirical[duplication.empirical > max(speciation.empirical)])` high duplication node ages eventually led to much larger expected variances for gene duplication events (median expected variance for duplication events preceding the oldest speciation events: `r round(median(Exp_var.duplication.precede.speciation),0)`, median expected variance for speciation events: `r round(median(Exp_var.speciation),0)`). This explains why the mean paralog distance was ~`r round((paralog_distance_mean/ortholog_distance_mean),1)` times higher than that of the mean ortholog distance (S4A Fig). This data distribution also makes it appear as if no speciation happened before `r max(speciation.empirical)` My, a problem shared with the pairwise analysis of the same data [@RN762]. There are also branch length issues for duplication events younger than the oldest speciation events, including very short branches and negative branch lengths. All this led to obtain abnormally high duplication contrasts for the empirical data. To limit such problems, Dunn et al. [@RN760] removed node contrasts higher than 0.5 on the empirical data (this does not impact Fig 1, as the simulation data never has such high contrasts). Following this practice, there are still higher expected variances following gene duplication events (S4A and S4B Figs). In the null simulations only the $\tau$ values are simulated, while the branch lengths (hence the variances) are taken from the empirical data, and thus share its biases. This explains why contrasts are lower for duplications than for speciations under null simulations as well as with empirical data.

\vspace{0.5cm}

__Randomization tests to assess the performance of phylogenetic method__

```{r randomization_tau_trees, eval=T, echo=F, message=F, warning=F}

  ## Since our dataset contains many speciation trees with no duplication, this may bias the result.
  ## So we identified calibrated trees with atleast one duplication events 
  trees_of_interest <- mclapply(gene_trees_pic1,tree_with_duplication, mc.cores = cores)
  trees_of_interest <- trees_of_interest[!is.na(trees_of_interest)]

  ## All calibrated trees with permuted actual Tau of the tips 
  Tau_randomized_tree <- mclapply(gene_trees_pic,shuffling_tau,mc.cores = cores)

  ## Permuted Tau for calibrated trees with atleast one duplication event
  Tau_randomized_filtered_tree <- mclapply(trees_of_interest,shuffling_tau,mc.cores = cores)

  ## Calculated PIC based on randomized Tau for both list of trees
  Tree_pic_randomized <- mclapply(Tau_randomized_tree,contrast_random,mc.cores = cores)
  Filtered_tree_pic_randomized_tau <- mclapply(Tau_randomized_filtered_tree,contrast_random,mc.cores = cores)

  ## Summarizing data of all trees in single dataframe, and removing rows for which contrasts are not available (e.g. for tips)
  summary_tree_pic_randomized <- bind_rows(lapply(Tree_pic_randomized,summary_function))
  summary_filtered_tree_pic_randomized_tau <- bind_rows(lapply(Filtered_tree_pic_randomized_tau,summary_function))
  Random_tau_contrast <- summary_tree_pic_randomized[which(!is.na(summary_tree_pic_randomized$pic_abs_random)),]
  Random_tau_contrast_filtered_tree <- summary_filtered_tree_pic_randomized_tau[which(!is.na(summary_filtered_tree_pic_randomized_tau$pic_abs_random)),]
  Random_tau_contrast <- Random_tau_contrast[which(Random_tau_contrast$pic_abs_random < 0.5),]
  Random_tau_contrast_filtered_tree <- Random_tau_contrast_filtered_tree[which(Random_tau_contrast_filtered_tree$pic_abs_random < 0.5),]

  ## Removing rows with NA event, and collecting all nodes contrast 
  nodes_contrast_random <- Random_tau_contrast[!is.na(Random_tau_contrast$D),]
  nodes_contrast_random_tau_filtered_tree <- Random_tau_contrast_filtered_tree[!is.na(Random_tau_contrast_filtered_tree$D),]
  nodes_contrast_random$Event <- factor(nodes_contrast_random$Event, levels=c("Speciation", "Duplication"))
  nodes_contrast_random_tau_filtered_tree$Event <- factor(nodes_contrast_random_tau_filtered_tree$Event, levels=c("Speciation", "Duplication"))

  ## Collecting PICs and expected variances of events after randomizing actual Tau 
  speciation_pic.r <- nodes_contrast_random$pic_abs_random[which(nodes_contrast_random$Event=="Speciation")]
  duplication_pic.r <- nodes_contrast_random$pic_abs_random[which(nodes_contrast_random$Event=="Duplication")]
  speciation_pic.var <- nodes_contrast_random$Variance[which(nodes_contrast_random$Event=="Speciation")]
  duplication_pic.var <- nodes_contrast_random$Variance[which(nodes_contrast_random$Event=="Duplication")]

  ## Performing test 
  wilcox_1_tailed.r <- one_tailed_wilcox(duplication_pic.r,speciation_pic.r)
  wilcox_2_tailed.r <- two_tailed_wilcox(duplication_pic.r,speciation_pic.r)
  variance_difference<- wilcox.test(duplication_pic.var,speciation_pic.var, alternative = "greater")$p.value
  if(variance_difference == 0) variance_diff <- 2.2e-16
  if(variance_difference!= 0) variance_diff <- variance_difference

  ## Median PICs of randomized Tau of the two events of all calibrated trees
  median_nodes_contrast_random <- NULL
  median_nodes_contrast_random <- aggregate(pic_abs_random~Event, nodes_contrast_random, median)
  median_nodes_contrast_random$pic.round <- paste0("Median = ",round(median_nodes_contrast_random$pic_abs_random,4),sep="")

  ## Median PICs of randomized Tau of the two events of trees with atleast one duplication
  median_nodes_contrast_random_tau_ftree <- NULL
  median_nodes_contrast_random_tau_ftree <- aggregate(pic_abs_random~Event, nodes_contrast_random_tau_filtered_tree, median)
  median_nodes_contrast_random_tau_ftree$pic.round <- paste0("Median = ",round(median_nodes_contrast_random_tau_ftree$pic_abs_random,4),sep="")

```

```{r randomization_events_trees, eval=T, echo=F, message=F, warning=F}

  ## All calibrated trees with permuted actual internal node events
  Event_randomized_pic_tree <- mclapply(gene_trees_pic, shuffling_event, mc.cores = cores)

  ## Permuting internal node events for trees with atleast one duplication event
  Event_randomized_filtered_tree <- mclapply(trees_of_interest,shuffling_event,mc.cores = cores)

  ##Summarizing data of all list of trees in a dataframe
  summary_event_randomized_pic_tree <- bind_rows(lapply(Event_randomized_pic_tree,summary_function))
  summary_filtered_tree_pic_randomized_event <- bind_rows(lapply(Event_randomized_filtered_tree,summary_function))

  Random_event_contrast <- summary_event_randomized_pic_tree[which(!is.na(summary_event_randomized_pic_tree$pic)),]
  Random_event_contrast_filtered_tree <- summary_filtered_tree_pic_randomized_event[which(!is.na(summary_filtered_tree_pic_randomized_event$pic)),]
  Random_event_contrast <- Random_event_contrast[which(Random_event_contrast$pic < 0.5),]
  Random_event_contrast_filtered_tree <- Random_event_contrast_filtered_tree[which(Random_event_contrast_filtered_tree$pic < 0.5),]

  ## Removing NA Events and collecting nodes contrasts of newly assigned events
  nodes_contrast_random_events <- Random_event_contrast[!is.na(Random_event_contrast$event_new),]
  nodes_contrast_random_event_filtered_tree <- Random_event_contrast_filtered_tree[!is.na(Random_event_contrast_filtered_tree$event_new),]
  nodes_contrast_random_events$event_new <- factor(nodes_contrast_random_events$event_new, levels=c("Speciation", "Duplication"))
  nodes_contrast_random_event_filtered_tree$event_new <- factor(nodes_contrast_random_event_filtered_tree$event_new, levels=c("Speciation", "Duplication"))

  ## Collecting PICs and expected variances of events after randomization 
  speciation_pic_revent <- abs(nodes_contrast_random_events$pic[which(nodes_contrast_random_events$event_new=="Speciation")])
  duplication_pic_revent <- abs(nodes_contrast_random_events$pic[which(nodes_contrast_random_events$event_new=="Duplication")])
  speciation_pic_rvar <- nodes_contrast_random_events$var_exp[which(nodes_contrast_random_events$event_new=="Speciation")]
  duplication_pic_rvar <- nodes_contrast_random_events$var_exp[which(nodes_contrast_random_events$event_new=="Duplication")]

  ## Performing test 
  wilcox_1_tailed_revent <- one_tailed_wilcox(duplication_pic_revent,speciation_pic_revent)
  wilcox_2_tailed_revent <- two_tailed_wilcox(duplication_pic_revent,speciation_pic_revent)

  ## Median PICs of randomized events of all calibrated trees 
  median_random_event <- NULL
  median_random_event <- aggregate(pic~event_new, nodes_contrast_random_events, median)
  median_random_event$pic.round <- paste0("Median = ",round(median_random_event$pic,4),sep="")

  ## Median PICs of randomized events of trees with atleast one duplication
  median_nodes_contrast_random_event_ftree <- NULL
  median_nodes_contrast_random_event_ftree <- aggregate(pic~event_new, nodes_contrast_random_event_filtered_tree, median)
  median_nodes_contrast_random_event_ftree$pic.round <- paste0("Median = ",round(median_nodes_contrast_random_event_ftree$pic,4),sep="")

```

```{r additional_analyses, eval=T, echo=F, message=F, warning=F}

  ## Collecting tree data (numner of tips, internal node events proportions) for all calibrated trees in a single dataframe
  tree_data_all <- bind_rows(lapply(gene_trees_pic,tree_data_collection))
  tree_data_plot <- na.omit(tree_data_all) ## Removing data for trees having no duplication events

  ## Dataframe for scatter plot
  data_scatterplot <- data.frame(ntip=NA,Expected_variance=NA,Event=NA)
  data_scatterplot <- rbind(data_scatterplot, data.frame(ntip=tree_data_plot$tip_num, Expected_variance=tree_data_plot$spe_var,Event="Speciation"))
  data_scatterplot <- rbind(data_scatterplot, data.frame(ntip=tree_data_plot$tip_num,Expected_variance=tree_data_plot$dup_var,Event="Duplication"))
  data_scatterplot <- data_scatterplot[-1,]
  data_scatterplot$Event <- factor(data_scatterplot$Event, levels=c("Speciation", "Duplication"))

  ## Plotting number of tips vs. expected variance of events
  event_variance <- ggplot(data_scatterplot,aes(x=ntip))+
    geom_line(aes(y=Expected_variance, color=Event),alpha=0.5)+
    xlab(expression(bold("Number of tips"))) +
    ylab(expression(bold("Expected variance"))) +
    xlim(0,180) +
    ylim(0,2000) +
    theme_classic() +
    theme(legend.title=element_blank(),legend.position=c(0.7,0.8)) +
    theme(axis.text=element_text(size=10,face="bold")) +
    ggtitle(expression(bold("B. Variance distribution for events")))+
    theme(plot.title = element_text(face="bold"))

  ## Creating dataframe for plotting events proportion distribution for all 8520 calibrated trees
  plot_data <- data.frame(proportion_event=NA,Event=NA)
  plot_data <- rbind(plot_data, data.frame(proportion_event=tree_data_all$spe_prop,Event="Speciation"))
  plot_data <- rbind(plot_data, data.frame(proportion_event=tree_data_all$dup_prop,Event="Duplication"))
  plot_data <- plot_data[-1,]
  plot_data$Event <- factor(plot_data$Event, levels=c("Speciation", "Duplication"))

  ## Histogram to show distribution of proportion of events for all calibrated trees
  event_histogram <- ggplot2_histogram_mod(data=plot_data, xName='proportion_event',
    groupName='Event',legendPosition="right",
    alpha=1, addDensity=TRUE,
    addMedianLine=TRUE, medianLineColor=c("#F8766D","#00BFC4"), medianLineSize=1) +
    theme_classic() +
    ggtitle(expression(bold("D. Event proportion plot")))+
    theme(plot.title = element_text(face="bold"))+
    xlab("Proportion of events") + 
    ylab("Density") +
    theme(legend.title=element_blank(),legend.position=c(0.9,0.8),legend.text=element_text(size=8)) +
    theme(axis.text=element_text(size=10,face="bold")) +
    theme(plot.title = element_text (face="bold", 
                              colour="black", lineheight=1.0),
          axis.title.x=element_text( face="bold",
                              colour="black", hjust=0.5),
          axis.title.y=element_text(face="bold",
                              colour="black", vjust=0.5, angle=90))

```

We used randomization tests to assess whether the results of different analyses of the empirical dataset are reliable and unbiased. In a first randomization test, we randomized the trait values (i.e. randomly permuting the $\tau$ values) across the tips of each tree without altering the node events of the trees. We then computed the contrasts for the speciation and duplication events of the trees when there is no relation between trait values and phylogenetic relationships. When we compared the nodes contrasts of speciation and duplication events of these randomized trees (Fig 2A), we found the same pattern as reported for the empirical gene trees by Dunn et al. [@RN760]. This shows that the PIC applied to these data is not measuring phenotypic evolution. It confirms that results are driven by their large differences in branch lengths (i.e. in expected variances) (Fig 2B), as on simulated null data. Any effect of trait divergence rates of speciation and duplication events is masked by this branch length difference. This violates the basic assumption of applicability of the PIC method to Brownian trait evolution. To remove the problem of difference in expected variances of the two events, we performed a second randomization test: we kept the original $\tau$ value for tips but randomly shuffled the events (duplication or speciation) of internal nodes of the empirical gene trees to maintain the original proportions of speciation and duplication events. The resulting trend (Fig 2C) still resembled the empirical gene trees data of Dunn et al. [@RN760]. This appears due to the fact that the majority of the tree events are speciation events (Fig 2D) with speciation node ages $\leq$ 296 My. Most of the trees with many duplication events on the other hand have ancient duplication events (duplication node age > 296 My) for which the evolutionary rates of duplication are often masked by the effect of longer branch lengths. Hence, opposite to our expectation, the calibrated trees with no or few duplications have higher overall nodes contrast (apparent fast evolution) than trees with many duplications (apparent slow evolution). This might be due to greater difficulty in detecting paralogs for fast evolving genes. Therefore, reshuffling of the events may not change the observed pattern of higher speciation contrasts than duplication contrasts.

Out of `r length(gene_trees_calibrated)` calibrated trees, `r trees_with_no_duplication` were species-only trees with no duplication events. For these `r trees_with_no_duplication` trees, random shuffling of events had no impact. To avoid this bias, we removed those `r trees_with_no_duplication` speciation trees as well as trees with negative branch lengths, and randomized the trait or the internal node events 100 times on the remaining `r length(trees_of_interest)` trees. However, we still always obtained significantly higher contrasts of speciation than of duplication (S5A and S5B Figs). The randomization tests pattern is the same when we used `r subset.trees` trees with strong phylogenetic signals (S5C and S5D Figs). All these analyses indicate that the results reported by Dunn et al. [@RN760] are biased by the available gene trees, and that this bias is not easy to correct.

```{r Fig2, dpi=400,fig.width=8, fig.height=8, fig.cap="**Fig 2**: Analyses on calibrated empirical gene trees of Dunn et al. [@RN760]. *P* values are from  Wilcoxon two-tailed tests. (A) Randomly shuffling the $\\tau$ values of the tips for 8520 gene trees does not alter the trend of empirical result [@RN760] showing an opposite trend to the ortholog conjecture. (B) The figure indicates the expected variance is much higher for duplication than speciation events irrespective of the number of tips considered for the study. (C) Using the original $\\tau$ data, if we permute the events (Speciation/Duplication/NA) of the nodes, the trend of result still resembles the empirical result obtained by the recent study [@RN760]. (D) The proportions of speciation events are much higher than duplication events for all time-calibrated trees. The dotted line represents the median proportion of both events. This plot shows a high frequency of trees have no/few duplication events."}

  ## Plotting the randomization data result for all calibrated trees 
  ## 1st plot
  jitterplot_randomized_tau <- 
    ggplot(nodes_contrast_random,aes(x=Event, y=pic_abs_random, fill=Event)) + 
    geom_jitter(aes(colour = Event),position=position_jitter(0.02), alpha=0.02) +
    geom_text(data = median_nodes_contrast_random, aes(label = pic.round),fontface = 2,size = 4, 
              hjust=0.5,vjust =-2.5)+
    geom_crossbar(data=median_nodes_contrast_random, aes(ymin = pic_abs_random, ymax = pic_abs_random),
                  size=0.2,col= "black", width = .2)+
    xlab( NULL ) +
    ylab(expression(bold("Phylogenetic Independent Contrasts (PICs)"))) +
    ylim(0, 0.4) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(1.9,1.9)) +
    theme(axis.text=element_text(size=10,face="bold")) +
    theme(legend.text=element_text(size=10,face="bold")) +
    annotate("text", x = 1.5, y = 0.3, label= wilcox_2_tailed.r, fontface = 4)+
    ggtitle(expression(bold(paste("A. Plot using randomized trait, ", tau))))+
    theme(plot.title = element_text(face="bold"))

  ## 2nd plot
  nodes_contrast_random_events$event_new <-factor(nodes_contrast_random_events$event_new, levels=c("Speciation", "Duplication"))

  jitterplot_randomized_event <- 
    ggplot(nodes_contrast_random_events,aes(event_new, y=pic, fill=event_new)) + 
    geom_jitter(aes(colour = event_new),position=position_jitter(0.02), alpha=0.02) +
    geom_text(data = median_random_event, aes(label = pic.round),fontface = 2,size = 4, 
              hjust=0.5,vjust = -2.5)+
    geom_crossbar(data=median_random_event, aes(ymin = pic, ymax = pic),
                  size=0.2, col= "black",width = .2)+
    xlab( NULL ) +
    ylab( NULL) +
    ylim(0, 0.4) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(1.9,1.9)) +
    theme(axis.text=element_text(size=10,face="bold")) +
    theme(legend.text=element_text(size=10,face="bold")) +
    annotate("text", x = 1.5, y = 0.3, label= wilcox_2_tailed_revent, fontface = 4)+
    ggtitle(expression(bold(paste("C. Plot using randomized events"))))+
    theme(plot.title = element_text(face="bold"))

  ##Plotting the result
  grid.arrange(jitterplot_randomized_tau,jitterplot_randomized_event, event_variance,event_histogram, nrow=2,ncol=2)

```

\vspace{0.5cm}
__The ortholog conjecture when $\tau$ evolution follows a Brownian model __

```{r Analysis_with_Caper_package_without_data_transformation, eval=T, echo=F, message=F, warning=F}

  ## In this section, we plan to perform diagnostic tests to check for Brownian model of Tau evolution
  ## This diagnostic test function takes a lot of time, so we previously ran "Premanuscript_run_TM.R" and saved the results
  ## Stored data are available in the following link https://doi.org/10.5281/zenodo.3354285 
  ## We loaded those stored results for further analyses
  load("/Users/admin/Desktop/nwork/simulation/Data_TMRR.rda")

  ## Summering results of different lists of BM trees (for details check "Premanuscript_run_TM.R" and S4 Fig)
  summary_brownian_tree <- bind_rows(lapply(standarized_pic_tree_with_dup_final,summary_function))
  summary_ancient_speciation_tree <- bind_rows(lapply(standarized_pic_tree_ancient_speciation,summary_function))
  Contrast_brownian <- summary_brownian_tree[which(!is.na(summary_brownian_tree$pic_Tau)),]
  Contrast_ancient_speciation_tree <- summary_ancient_speciation_tree[which(!is.na(summary_ancient_speciation_tree$pic_Tau)),]

  ## We have checked for maximum PIC for all sets of Brownian trees, and the PIC was always less than 0.5

```

```{r Fig3, dpi=400,fig.width=8, fig.height=4, fig.cap="**Fig 3**: The ortholog conjecture test on $\\tau$ for trees passing diagnostic plot tests with proper controls. *P* values are from Wilcoxon two-tailed tests. 'PICs' refers 'Phylogenetic Independent Contrasts'. Contrasts of speciation and duplication events reject the null hypothesis with different node age limits for the trees following Brownian model of $\\tau$ evolution, and provide support for the ortholog conjecture."}

  ## Analysis considering duplication age <= oldest speciation age
  speciation_pic_nodeage <- Contrast_brownian$node_age[which(Contrast_brownian$Event=="Speciation")]
  Contrast_brownian_plot <- Contrast_brownian[which(Contrast_brownian$node_age <= max(speciation_pic_nodeage)),]

  # Considering the relevant columns for plotting and removing NA events
  Contrast_brownian_plot <- na.omit(Contrast_brownian_plot[c(6,18:19,22:24)])
  #Contrast_brownian_plot$pic_Tau<-abs(Contrast_brownian_plot$pic_Tau) ## Taking the absolute value
  Contrast_brownian_plot$Events <- factor(Contrast_brownian_plot$Event, levels=c("Speciation", "Duplication"))

  # Extracting speciation and duplication PICs  
  speciation_contrast_tau <-  abs(Contrast_brownian_plot$pic_Tau[which(Contrast_brownian_plot$Events=="Speciation")])
  duplication_contrast_tau <-  abs(Contrast_brownian_plot$pic_Tau[which(Contrast_brownian_plot$Events=="Duplication")])

  ## Performing test 
  wilcox_test_tau <- two_tailed_wilcox(duplication_contrast_tau,speciation_contrast_tau)
  title <- paste0("Plots on contrasts standardized ",length(standarized_pic_tree_with_dup_final), " gene trees", sep="")

  ## 2nd analysis considering duplication age is 25% higher than oldest speciation age
  nodeage_limit <- max(speciation_pic_nodeage)+ 0.25*max(speciation_pic_nodeage)
  Contrast_brownian_plot2 <- Contrast_brownian[which(Contrast_brownian$node_age <= nodeage_limit),]
  Contrast_brownian_plot2 <- na.omit(Contrast_brownian_plot2[c(6,18:19,22:24)])
  Contrast_brownian_plot2$Events <- factor(Contrast_brownian_plot2$Event, levels=c("Speciation", "Duplication"))

  # Separating speciation and duplication contrast data
  speciation_contrast_tau2 <-  abs(Contrast_brownian_plot2$pic_Tau[which(Contrast_brownian_plot2$Events=="Speciation")])
  duplication_contrast_tau2 <-  abs(Contrast_brownian_plot2$pic_Tau[which(Contrast_brownian_plot2$Events=="Duplication")])

  ## Performing test 
  wilcox_test_tau2 <- two_tailed_wilcox(duplication_contrast_tau2,speciation_contrast_tau2)

  ## Making a  dataframe for plotting
  plot_df <- data.frame(pic=NA,var=NA, group=NA, Event=NA) ## declaring data frame
  plot_df <- rbind(plot_df, data.frame(pic=abs(Contrast_brownian_plot$pic_Tau), var= Contrast_brownian_plot$variance, group="Age <= 296 My", Event=Contrast_brownian_plot$Events))
  plot_df <- rbind(plot_df, data.frame(pic=abs(Contrast_brownian_plot2$pic_Tau), var= Contrast_brownian_plot2$variance, group="Age <= 370 My", Event=Contrast_brownian_plot2$Events))
  plot_df <- plot_df[-1,]
  plot_df$Event <- factor(plot_df$Event, levels=c("Speciation", "Duplication"))

  ## Boxplot
  dodge <- position_dodge(width = 0.51)

    ggplot(plot_df, aes(group, pic, fill = Event)) +
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    #geom_jitter(aes(colour = Event),position=dodge, alpha=0.02) +
    xlab(NULL) +
    ylab(expression(bold(paste("PICs of ",tau)))) +
    coord_cartesian(ylim=c(0, 0.06)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9)) +
    #theme(legend.position="none")+
    theme(axis.text.y = element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold")) +
    ggtitle(paste0(title))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x=1.04, y=0.05, label=wilcox_test_tau, fontface=4) +
    annotate("text", x=1.98, y=0.05, label=wilcox_test_tau2, fontface=4)

```

Following a Brownian model (BM), the small and large values of standardized contrasts should be equally as likely to occur on any node of the tree [@RN783]. Diagnostic plot tests (details in the Methods) for each tree can indicate whether trait evolution follows BM for that tree [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816]. Since BM is intrinsic to the implementation of PIC method, we identified trees that passed diagnostic plot tests at 95% confidence level for $\tau$. Hence, we used a subset of `r length(standarized_pic_tree_with_dup_final)` gene trees by removing trees with negative branch lengths, pure speciation trees, and trees that failed diagnostic tests for BM (S6 Fig). Moreover, to meet the correct branch length assumption of PIC applicability on trees with gene duplications, we used limitations of node age on gene duplication. It is necessary to put such a limitation to avoid biases due to inaccurate calibration of very old duplication nodes. Analyses on `r length(standarized_pic_tree_with_dup_final)` these trees with a node age limit of `r max(speciation_pic_nodeage)` My (i.e. node age $\leq$ maximum speciation age), provided support for the ortholog conjecture (Fig 3), in contrast to previous results. We repeated the analysis including slightly older duplication events (duplication age $\leq$ `r round((max(speciation_pic_nodeage)+ 0.25*(max(speciation_pic_nodeage))),0)` My, i.e. maximum 25% older than the oldest speciation). We still were able to detect higher contrasts for duplication than for speciation (Fig 3). Finally, randomization tests on these trees indicated no bias in the inference (S7A-S7D Figs), supporting the relevance of the inferences on empirical data.

Among these `r length(standarized_pic_tree_with_dup_final)` trees (S6 Fig), very old duplication node contrasts were also involved to identify contrast distribution of $\tau$ as per BM, which might bias the analysis. Hence, we repeated the identification of BM trees by standard diagnostic plot tests considering only `r length(trees_of_interest)` trees with at least one duplication event, after removing very old duplication nodes (duplication age > `r round((max(speciation_pic_nodeage)+ 0.25*(max(speciation_pic_nodeage))),0)` My) (S6 Fig). We still were able to find support for the ortholog conjecture on these identified `r length(standarized_pic_tree_with_dup_final_new)` BM trees (S8 Fig).

To avoid calibration bias of older duplication events, we also considered a more restricted set of `r length(standarized_pic_tree_ancient_speciation)` passing more strict conditions (S6 Fig), i.e. BM trees in which there is at least one duplication, and a speciation event older than all duplications. Calibration bias for duplication events is expected to be very limited for such trees. Support for the ortholog conjecture still holds with these `r length(standarized_pic_tree_ancient_speciation)` trees (Fig 4A). Randomization tests on these `r length(standarized_pic_tree_ancient_speciation)` trees also support the biological relevance of these results (Figs 4B and 4C,S9 Fig). These observations confirm that the ortholog conjecture is supported for adequately standardized BM trees.

```{r reanalyses_Dunn_data_BM_OU,echo=F, message=F, warning=F}

  ## Testing the ortholog cojecture with nodes contrasts of 3151 Brownian trees provided by Dunn et al.
  speciation_BM_pic <- nodes_contrast_filtered_bm$pic[which(nodes_contrast_filtered_bm$Event=="Speciation")]
  duplication_BM_pic <- nodes_contrast_filtered_bm$pic[which(nodes_contrast_filtered_bm$Event=="Duplication")]

  ## Performing test 
  wilcox_one_tailed_BM <- wilcox.test(duplication_BM_pic,speciation_BM_pic,alternative = "greater")$p.value
  wilcox_two_tailed_BM <- wilcox.test(duplication_BM_pic,speciation_BM_pic)$p.value

  ## Identifying the index of BM trees of Dunn et al. from their all time calibrated trees
  ## and extracting them based on their criteria for further use
  tree_summary$index <- as.numeric(rownames(tree_summary))
  genes_pass_BM <- tree_summary$index[tree_summary$dAIC_bm < 2]
  BM_trees_model_fit <- gene_trees_calibrated[c(genes_pass_BM)]

```

Instead of using standard diagnostic plot tests [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816], Dunn et al. [@RN760] considered model-fitting criteria to identify BM trees from their list of `r length(gene_trees_calibrated)` calibrated trees. They contrasted the BM model (neutral drift model with no selection) to the Ornstein-Uhlenbeck (OU) model (i.e. an extended BM model with stabilizing selection) [@RN843;@RN841] for this purpose. After computing the difference in Akaike Information Criterion ($\Delta$AIC) [@RN380] scores of the models for each tree, they identified 3151 BM trees, out of which only 8 trees purely favored the BM model ($\Delta$AIC~BM~ < 2 and $\Delta$AIC~OU~ $\ge$ 2). The other 3143 trees favored both the BM and OU models ($\Delta$AIC~BM~ < 2 and $\Delta$AIC~OU~ < 2). A Wilcoxon two-tailed test on those 3151 trees provided similar results to the `r length(gene_trees_calibrated)` calibrated trees, i.e. higher $\tau$ evolution for orthologs than for paralogs (PIC~speciation~ = `r round(median(speciation_BM_pic),4)`, PIC~duplication~ = `r round(median(duplication_BM_pic), 4)`, *P* = $`r format(wilcox_two_tailed_BM, digits= 3, scientific = TRUE)`$). This result implies that selection of BM or non-BM trees does not explain these results. A similar trend was observed for the recent duplicates (node age $\le$ `r max(speciation_pic_nodeage)` My) (S9A Fig). The pattern of systematic increase of duplication PICs with node age (S10A Fig) plausibly implies that the evolution of $\tau$ does not in fact follow Brownian motion for these trees. Randomization tests on those trees (S10B and S10C Figs) confirmed that contrasts were not appropriately standardized as per BM, and thus the inference drawn on empirical data was not supported. When we performed diagnostic plot tests on those 3151 trees, we found support for the ortholog conjecture for `r length(BM_tree_diagnostic_test_final)` trees passing the diagnostic tests (S11 Fig). All these analyses suggest that Dunn et al. did not identify accurately BM trees, and that such misidentification leads to violating the assumptions of phylogenetic comparative methods, and to erroneous results.

```{r Analyses_on_standardized_806_trees, eval=T, echo=F, message=F, warning=F}

  ## Considering the relevant columns and removing the NA data
  Contrast_brownian_806_trees <- na.omit(Contrast_ancient_speciation_tree[c(6,18:19,22:24)])
  Contrast_brownian_806_trees$Event <- factor(Contrast_brownian_806_trees$Event, levels=c("Speciation", "Duplication"))

  ## Taking the absolute values of PIC of Tau
  Contrast_brownian_806_trees$pic_Tau <- abs(Contrast_brownian_806_trees$pic_Tau)

  ## Extracting speciation and duplication PICs for these 806 trees passing diagnostic plots and having ancient speciation event
  speciation_pic_tau_806 <- speciation_contrast(Contrast_brownian_806_trees,'pic_Tau')
  duplication_pic_tau_806 <- duplication_contrast(Contrast_brownian_806_trees,'pic_Tau')

  ## Median value collection for Tau of the standarized trees
  median_tau_806 <- median_jitter((Contrast_brownian_806_trees),'pic_Tau','Event')
  colnames(median_tau_806)[1] <- 'Event'
  colnames(median_tau_806)[2] <- 'pic_Tau'

  ## Performing test 
  p_tau_806 <- paste0("Two-tailed test, ", two_tailed_wilcox(duplication_pic_tau_806,speciation_pic_tau_806))

```

```{r Additional_analyses_on_806_BM_trees, eval=T, echo=F, message=F, warning=F}

  ## Collecting tree data for all calibrated trees in a single dataframe
  tree_data_806<-bind_rows(lapply(standarized_pic_tree_ancient_speciation,tree_data_collection))
  median_number_of_tips<-median(tree_data_806$tip_num)
  median_number_of_events<-median(tree_data_806$internal_events)
  median_proporion_of_speciation<-median(tree_data_806$spe_prop)
  median_proporion_of_duplication<-median(tree_data_806$dup_prop)
  median_proporion_of_NA_event<-median(tree_data_806$NA_prop)

```

```{r Fig4, dpi=400,fig.width=8, fig.height=8, fig.cap= paste0("**Fig 4**: Results on real and randomized data for ", length(standarized_pic_tree_ancient_speciation), " BM trees with ancient speciation events. *P* values are from  two-tailed Wilcoxon tests. (A) The jitter plot on empirical gene tree data supports the ortholog conjecture. Main plots in (B) and (C) show the distributions of two-tailed test *P* values estimated for the whole set of ", length(standarized_pic_tree_ancient_speciation), " BM trees by permuting $\\tau$ (B), and by permuting the internal node events (C) ('Speciation', 'Duplication', and 'NA') for 1000 independent runs respectively. Inset plots in (B) and (C) show *P* value distributions plots of one-tailed test to check if there was an excess of duplication contrasts over speciation contrasts by randomizing $\\tau$, and events respectively. The proportions of speciation (median = ", median_proporion_of_speciation, "), and NA events (median = ", median_proporion_of_NA_event, ") are much higher than that of duplication events (median = ", median_proporion_of_duplication, ") in these trees (S9 Fig). Hence, higher duplication contrasts are more likely to be replaced by speciation events by permutations of events. Due to this reason, we found significant shifts towards left in the main plot of C, and towards *P* value 1 in the inset plot of C to depict larger contrasts for speciation events. Comparisons of plot A to plot B, and plot A to plot C show that randomization results in both the cases actually differ from the empirical result.")}

  ## Jitter plot with empirical data
  plot4AM <- jitter_plot_tau(Contrast_brownian_806_trees,median_tau_806,p_tau_806)+
            ylab(expression(bold(paste("PIC of ",tau))))+
            ggtitle(expression(bold(paste("A. Empirical result"))))

  ## Dataframe of P values from randomization tests on 806 trees
  pval_dataframe_806 <- data.frame(pval_tau_2sided=NA,pval_tau_1sided=NA,pval_event_2sided=NA,pval_event_1sided=NA,tree_set=NA)
  pval_dataframe_806 <- rbind(pval_dataframe_806, data.frame(pval_tau_2sided=pval,pval_tau_1sided=pval_1sided,pval_event_2sided=pval_event,pval_event_1sided=pval_event_1sided,tree_set="806_tree"))
  pval_dataframe_806 <- pval_dataframe_806[-1,]

  ## Plots on P value distribution using randomized Tau
  plot4BM <- histo_plot(pval_dataframe_806,'pval_tau_2sided','tree_set') +   
            xlab(expression(bold(paste( bolditalic(P)," value (two-tailed test)")))) 
  plot4BM <- plot4BM + 
            ylim(0,150) +
            ggtitle(expression(bold(paste("B. Randomized ", tau))))+
            theme(plot.title = element_text(face="bold"))+
            theme(axis.text=element_text(size=10,face="bold"))
  plot4CM <- histo_plot(pval_dataframe_806,'pval_tau_1sided','tree_set') +   
            xlab(expression(bold(paste( bolditalic(P)," value (one-tailed test)")))) 
  plot4CM <- plot4CM + 
            ylim(0,150) + 
            #ylab(" ")+ggtitle(expression(bold(paste("C. Owo sided tests on randomized ", tau))))+
            theme(plot.title = element_text(face="bold"))+
            theme(axis.text=element_text(size=10,face="bold"))

  ## Plots on P value distribution using randomized internal node events
  plot4DM <- histo_plot(pval_dataframe_806,'pval_event_2sided','tree_set') +
           xlab(expression(bold(paste( bolditalic(P)," value (two-tailed test)"))))
  plot4DM <- plot4DM + 
           ylim(0,150) +
           ggtitle(expression(bold(paste("C. Randomized events"))))+
           theme(plot.title = element_text(face="bold"))+
           theme(axis.text=element_text(size=10,face="bold"))
  plot4EM <- histo_plot(pval_dataframe_806,'pval_event_1sided','tree_set') +
           xlab(expression(bold(paste( bolditalic(P)," value (one-tailed test)"))))
  plot4EM <- plot4EM + 
           ylim(0,150) +
           #ylab(" ")+
           #ggtitle(expression(bold(paste("E. One sided tests on randomized events"))))+
           theme(plot.title = element_text(face="bold"))+
           theme(axis.text=element_text(size=10,face="bold"))

  ## Preparing final plots and arranging them
  plot4B <- ggdraw() +
    draw_plot(plot4BM)+
    draw_plot(plot4CM, x = .5, y = .5, width = .5, height = .5)

  plot4C <- ggdraw() +
    draw_plot(plot4DM)+
    draw_plot(plot4EM, x = .5, y = .5, width = .5, height = .5)

  # grid.arrange(plot4AM,grid.arrange(plot4BM,plot4CM,plot4DM,plot4EM,nrow = 2, ncol=2)) 
  gridExtra::grid.arrange(plot4AM,plot4B,plot4C,nrow = 3)

```

### Discussion

We agree with Dunn et al. [@RN760] that evolutionary comparisons should be done considering a phylogenetic framework when possible. However, this does not imply that phylogenetic independent contrasts can be applied easily to phylogenomics data and questions. To get a clear picture, we limited our study to the same phylogenetic gene trees used by Dunn et al. [@RN760]. Our reanalysis identified problems generated during the time calibration of duplication nodes of pruned trees for which no external time references are available, and with non respect of the hypothesis of Brownian motion. The strongest bias was for duplication nodes preceding the oldest speciation nodes. This, in turn, introduced several biases in their analyses, and influenced their results. When we identified and controlled for such biases, PIC results changed to support the ortholog conjecture, consistent with our previous pairwise approach [@RN762] on the same $\tau$ data.

To our knowledge, no one before Dunn et al. [@RN760] applied a phylogenetic approach in comparative biology to study the effect of gene duplication on functional evolution, despite an early call to do so [@RN837]. In line with previous studies [@RN782;@RN851], we explored whether the application of a phylogenetic method might inflate errors (e.g. rejection of the null hypothesis in null condition, Figs 1A and 1B) if applied without thorough testing for the fundamental assumptions of the method. Assumptions of proper branch length information and of Brownian model of trait evolution are related, so that modifications of branch lengths can change the evolutionary model [@RN851]. We find that branch length calibration is mostly inaccurate for old duplication events (duplication events prior to the oldest speciation event in the data) (S2 Fig). Use of such node contrasts causes a strong rise in expected variance for duplication events compared to the speciation events (S4 Fig). This may bring about lack of statistical power to detect the signal of ortholog conjecture, when in reality the signal is present, and even bias towards a pseudo-signal. As a remedial measure, we limited our analysis to the trees for which Brownian motion of trait evolution has been identified with the aid of standard diagnostic plot tests [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816]. To remove issues with branch length inaccuracies for older duplication events, we used limitation of node ages for duplication events. Using all such measures to control for biases, we found support for the ortholog conjecture (Figs 3 and 4A). The reliability of our inference was validated by two different randomization tests, which confirmed the fact that our result was not due to biases in the data or analysis (S7A-S7D Figs, Figs 4B and 4C).

Being aware of the fact that branch lengths are fundamental to phylogenetic comparative methods, Dunn et al. [@RN760] added random noise in the speciation calibration time point, and extended terminal branch lengths to test sensitivity of their observations. While extending terminal branch can reduce error rate due to the presence of negative branch lengths in `r length(trees_negative)` trees, none of the modifications appear to avoid biases when there are so vast differences in branch lengths [@RN853;@RN854;@RN855;@RN856;@RN852;@RN858;@RN859]. These can also lead to violating the Brownian model of trait evolution [@RN851]. Diagnostic plot tests appear necessary to assess adequate standardization of contrasts as per BM in such cases [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816].

The importance of proper BM tree selection should not be underestimated as it can impact the contrast analysis. There exist no fixed protocols to select BM trees. We used standard diagnostic plots recommended in earlier studies for PIC   [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816;@RN813;@RN825;@Rcore] to rule out significant departure from BM for each tree individually at 95% confidence level. In contrast, Dunn et al. [@RN760] used both BM and OU as null models for each tree, leading them to identify "BM trees" for which the two models were equally good fits. These do not appear to be proper BM trees according to several observations, since duplication contrasts increased with node age (S10A Fig). Moreover, the inferences based on such trees with empirical data were not different from that based on randomized trees (S10B and S10C Figs). When we applied diagnostic tests on those 3151 "BM trees", `r (length(BM_trees_model_fit)-length(BM_tree_diagnostic_test_final))` trees failed the diagnostic test, and rest of the trees provided support for the ortholog conjecture (S11 Fig).

Empirical support for the ortholog conjecture has been mixed, with most studies supporting, and a few failing to do so  [@RN770;@RN790;@RN793;@RN794;@RN762]), our results provided support for the ortholog conjecture using large-scale genome wide tissue specificity data in a phylogenetic framework after controlling for bias. Due to lack of detailed functional information, many studies are still limited to gene expression data as a proxy of function. Recently, using functional replaceability assay, experimental studies [@RN857;@RN861] have shown that orthologous genes can be swapped between essential yeast genes and human, although this is rarely the case for all the members of expanded human gene families [@RN861], validating one prediction of the ortholog conjecture.

These analyses are mainly based on gene trees dominated by small scale duplication events. In these trees, the age of the same duplication clade is never fixed. Use of time calibrated whole genome duplication trees could be beneficial in this regard. Since the approximate time period of whole genome duplications are known, we can test the hypothesis more conveniently by avoiding use of any node age limits on whole genome duplication events for those trees.

\pagebreak

### Methods

__Resource details for reanalyses__

Our reanalyses are based on 8520 annotated (with the events: speciation, duplication and NA), pruned, and time calibrated ENSEMBL Compara v.75 [@RN806] gene trees, having at least 4 tips with non null trait data.  They were obtained from Dunn *et al* [@RN760]. Like them [@RN760], we used a subset of precomputed $\tau$ data (as trait) of 8 vertebrate species from the study of Kryuchkova-Mostacci and Robinson-Rechavi [@RN762]. Kryuchkova-Mostacci and Robinson-Rechavi [@RN762] computed $\tau$ and mean gene expression levels by following the method of Yanai et al. [@RN215]. The subset of data was based on the RNA-seq data provided by Brawand et al. [@RN838] for 6 tissues. We used the same random seed number as in [@RN760] to reproduce the simulation results for reanalysis. All reproduced data of Dunn et al. were stored in the "manuscript_dunn.RData" file (https://doi.org/10.5281/zenodo.3354285). We used the results stored in the 'data' or 'phylo' slot of the trees for further analyses. To differentiate our own function from theirs [@RN760], we remodified the original function script of Dunn et al. from "functions.R" to "functions_Dunn.R". Some of the analyses were time consuming, so we made a separate script "Premanuscript_run_TM.R" to run before knitting the markdown file. We stored the outputs in ???Data_TMRR.rda??? file (https://doi.org/10.5281/zenodo.3354285) and loaded it during our analyses. All the details of different functions were provided inside the script. We supply all the previously stored data (to reduce computation time during reproduction of result) and function files including our own ("function_TM_new.R") with this manuscript. All scripts are available on GitHub: https://github.com/tbegum/Testing_the_ortholog_conjecture.

We used phylosig function() of "phytools" package [@RN804] to identify trees with phylogenetic signal (*P* < 0.05) using Blomberg's K [@RN819;@RN804;@RN800].

Analyses and plotting were performed in R version 3.5.1 [@Rcore] using treeio [@RN823], geiger [@RN796], stringr [@RN900],  digest[@RN909], dplyr [@RN908], tidyverse [@RN904], ggrepel[@RN910], gtools [@RN902], ggplot2 [@RN903], cowplot [@RN905], easyGgplot2 [@RN901], gridExtra [@RN906], and png [@RN907] libraries.

__Randomization test of $\tau$ values__

For each tree, we used $\tau$ data (column name "Tau" in each tree 'data' object) across the tips to carry out our randomization test. To randomize we permuted the actual $\tau$ data  without altering internal node events. The pic() function of the "ape" package [@RN803] was used to compute PIC of nodes  for each tree using permuted $\tau$ of tips. For each run, we compared the contrasts of speciation and duplication events of the whole set of randomized trees to estimate difference in event contrasts based on Wilcoxon signed rank test. For 100 or 1000 runs, we repeated the above process 100 or 1000 times to obtain a distribution plot of 100 or 1000 independent *P* values.

__Randomization test of node events__

Some of the speciation nodes had daughters with same clade names in the gene trees we used for our study. Dunn et al. changed such node events to "NA" to avoid problems during time calibration of the trees. Such annotated node event information ("Speciation", "Duplication", "NA") for each tree was available as "Event" in the tree 'data' slot. To randomize, we permuted the internal node events (added as column name "event_new" in the 'data' slot) by maintaining the actual proportion of events for each tree. Then, we used the PIC of actual $\tau$ at tips to estimate contrasts difference between newly assigned speciation and duplication node events by Wilcoxon rank tests. For 100 or 1000 independent runs, we repeated the same procedure to obtain 100 or 1000 independent *P* values.

__Checking for Brownian model of trait evolution__

There exists no established protocol to identify the BM (Brownian model) trees. Dunn et al. used model-fitting with $\Delta$AIC~BM~ = AIC~BM~ - min(AIC~BM~,AIC~OU~) and $\Delta$AIC~OU~ = AIC~OU~ - min(AIC~BM~,AIC~OU~). They applied $\Delta$AIC~BM~ < 2 as the relative support for the BM model. We used several diagnostic tests to test BM behavior, as recommended in several studies [@RN851;@RN779;@RN781;@RN782;@RN780;@RN783]. Among diagnostic tests, the most usual method for contrasts standardization is to check a correlation between the absolute values of standardized contrasts and their expected standard deviations [@RN816;@RN781;@RN782]. Under Brownian motion, there should be no correlation. This test and the correlation between the absolute values of standardized contrasts and the logarithm of their node age are model diagnostic plot tests in the caper ("Comparative Analyses of Phylogenetics and Evolution in R") package [@RN781;@RN813;@RN825; @Rcore]. We used both of them in our study by using the "crunch" algorithm of the "caper" package in R, which implements the methods originally provided in CAIC [@RN781; @RN813;@RN825; @Rcore]. Correlation of node heights with absolute values of contrasts has also been reported to be a reliable indicator of deviation from the Brownian model [@RN783]. Hence, we computed node height for each node in a tree using the ape package [@RN803]. We also used the correlations of node height and node depth to the absolute value of nodes contrasts to rule out significant trend in any of the 4 tests. We used *P* < 0.05 to assess a significant correlation for the diagnostic plot tests. A significant trend (positive or negative) reliably indicates a deviation from the BM of trait evolution for that tree [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816], and we removed those trees from our analysis.  Contrast calculation on negative branch lengths is not desirable, so we removed trees with negative branch lengths before applying the crunch() function. Trees passing all 4 diagnostic tests were considered as BM trees for our study.

### Acknowledgements

We sincerely acknowledge Martha Liliano Serranno Serranno for initial help in understanding the phylogenetic independent contrast method. We thank Nicolas Salamin, Julien Wollbrett, Jialin Liu, Sebastien Moretti, Sara Fonseca Costa, Kamil Jaron and all the members of the Robinson-Rechavi group for their help and useful discussions. We also acknowledge Cassey Dunn for his initial help in reproducing their results. Parts of the computations were performed at the Vital-IT (http://www.vital-it.ch) Center for high-performance computing of the SIB Swiss Institute of Bioinformatics.

\pagebreak

## Supporting Information

```{r Analysis_all_trees_with_different_random_seed_number,eval=T, echo=F, message=F, warning=F}

 treedata <- function(tree)
 {
  if(class(tree) == "treedata") 
  {
    tree@data <- tidytree::as_tibble(tree@data)
  }
  ## Reading tree data slot
  return(tree)
 }

  ## Applying the function on calibrated trees
  gene_trees_calibrated_new <- mclapply(gene_trees_calibrated,treedata, mc.cores = cores)

  ## Null simulation using Dunn et al. function
  ## Changing random seed number
  set.seed(123)

  gene_trees_sim_null_8520 <- foreach( tree=gene_trees_calibrated_new ) %dopar% 
sim_tau( tree )
  gene_trees_sim_null_8520 %<>% add_pics_to_trees()

  ## OC simulation using Dunn et al. function
  set.seed(123)
  gene_trees_sim_oc_8520 <- foreach( tree=gene_trees_calibrated_new ) %dopar% 
sim_tau( tree, dup_adjust=2)
  gene_trees_sim_oc_8520 %<>% add_pics_to_trees()

	## Making a dataframe by combining data slots of all these 8520 trees
	sim_null_contrast_8520_new <- bind_rows(lapply(gene_trees_sim_null_8520,summary_function))
	sim_oc_contrast_8520_new <- bind_rows(lapply(gene_trees_sim_oc_8520,summary_function))

	## Extracting the columns of event and PIC data from null and OC simulation result of trees with phylogenetic signals
  null_data_8520_new <- na.omit(sim_null_contrast_8520_new[c(6,20)])
  null_data_8520_new$label <- "Null simulation"
  OC_data_8520_new <- na.omit(sim_oc_contrast_8520_new[c(6,20)])
  OC_data_8520_new$label <- "OC simulation"

  ## Extracting speciation and duplication PICs from null and OC simulation of all calibrated trees
  speciation_null_8520 <- null_data_8520_new$pic[which(null_data_8520_new$Event=="Speciation")]
  duplication_null_8520 <- null_data_8520_new$pic[which(null_data_8520_new$Event=="Duplication")]
  speciation_oc_8520 <- OC_data_8520_new$pic[which(OC_data_8520_new$Event=="Speciation")]
  duplication_oc_8520 <- OC_data_8520_new$pic[which(OC_data_8520_new$Event=="Duplication")]

  ## Performing Wilcoxon tests on null and OC simulation data of all calibrated trees
  wilcox_2_tailed_null_8520_new <- two_tailed_wilcox(speciation_null_8520,duplication_null_8520)
  wilcox_2_tailed_oc_8520_new<- two_tailed_wilcox(speciation_oc_8520,duplication_oc_8520)

```

```{r Fig_S1, dpi=400,fig.width=8, fig.height=7, fig.cap= paste0("S1 Fig: Repeating simulations on all calibrated trees with different random seed number. *P* values are from Wilcoxon two-tailed tests. Simulations with different seed number did not change the trend of results as observed in Fig 1A.")}

  ## Dataframe is built based on 2896 trees with phylogenetic signals
  simulation_8520_new <- rbind(null_data_8520_new, OC_data_8520_new)

  ### Generating plot on simulation data for all calibrated trees 
  dodge <- position_dodge(width = 0.51)
  ggplot(simulation_8520_new,aes(x=label, y=pic, fill=Event)) + 
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    xlab(NULL) +
    ylab(expression(bold("Phylogenetic Independent Contrasts (PICs)"))) +
    coord_cartesian(ylim=c(0, 0.05)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9),legend.text=element_text(size=8)) +
    theme(axis.text.y= element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold"))+
    ggtitle(paste0("Plot on 8520 calibrated trees"))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x = 1.04, y = 0.035, label= wilcox_2_tailed_null_8520_new, fontface=4) +
    annotate("text", x = 1.98, y = 0.035, label= wilcox_2_tailed_oc_8520_new, fontface=4)

```

```{r Fig_S2, dpi=400,fig.width=8, fig.height=6, fig.cap= paste0("S2 Fig: An example of a calibrated gene trees used by Dunn et al. in their study. The blue nodes represent the speciation nodes while the red nodes are duplication nodes. For nodes with no number the event is assigned as NA by the authors [@RN760]. Since, the duplication time is unknown, the tree is calibrated based on speciation node ages. Notice that this gives an unreasonably large estimated date for the ancient duplication, resulting in higher expected variance and abnormally low contrast for the corresponding node event. This example is to demonstrate that it is often risky to use such nodes to infer the result of phylogenetic method that intends to study the effect of gene duplication.")}

  ## One example tree
  tree <- gene_trees_pic[[247]]
  ntips <- length(tree@phylo$tip.label) ## Number of tips for the tree

  ##Identifying internal speciation and duplication nodes from the 'tree@data' object
  nodes_duplication <- tree@data$node[which(tree@data$Event=="Duplication")]
  nodes_speciation <- tree@data$node[which(tree@data$Event=="Speciation" & tree@data$node > ntips)]

  ## Plotting the tree
  plotTree(tree@phylo, mar=c(5,2,2,2))
  axisPhylo()
  nodelabels(node=nodes_speciation,frame = "none", col="blue", font=2)
  nodelabels(node=nodes_duplication,frame = "none", col="red", font=2)
  mtext("An example of time calibrated trees for two different events",font=2,side=3, line=-1, outer = TRUE)
  mtext("Million Years (My)",font=2,side=1, line=-2, outer = TRUE)

```

```{r More_analysis_trees_with_strong_phylogenetic_signal,eval=T, echo=F, message=F, warning=F}

  ## Removing null and OC simulation trees associated with negative branch lengths
	gene_trees_pic1_sim_null <- gene_trees_sim_null[-c(trees_negative)]
  gene_trees_pic1_sim_oc <- gene_trees_sim_oc[-c(trees_negative)]

  ## Index number of trees with strong phylogenetic signal
  index_signal_tree <- signal_tree$tree_number[which(signal_tree$Pvalue_Blomberg< 0.05 & signal_tree$K_effecient > 0.551)]

  ## Collecting null and OC simulation trees associated with phylogenetic signals (P <0.05)
  tree_sim_null_1135 <- gene_trees_pic1_sim_null[c(index_signal_tree)]
  tree_sim_oc_1135 <- gene_trees_pic1_sim_oc[c(index_signal_tree)]

  #gene_trees_sim_null_test <- foreach( tree=tree_sim_null_1135 ) %dopar% sim_tau( tree )
  #gene_trees_sim_null_test %<>% add_pics_to_trees()

	## Making a dataframe by combining data slots of all these 1135 trees
	sim_null_contrast_1135 <- bind_rows(lapply(tree_sim_null_1135,summary_function))
	sim_oc_contrast_1135 <- bind_rows(lapply(tree_sim_oc_1135,summary_function))

	## Extracting the columns of event and PIC data from null and OC simulation result of trees with phylogenetic signals
  null_data.1135 <- na.omit(sim_null_contrast_1135[c(6,20)])
  null_data.1135$label <- "Null simulation"
  OC_data.1135 <- na.omit(sim_oc_contrast_1135[c(6,20)])
  OC_data.1135$label <- "OC simulation"

  ## Extracting speciation and duplication PICs from null and OC simulation of trees with phylogenetic signals
  speciation_null_1135 <- null_data.1135$pic[which(null_data.1135$Event=="Speciation")]
  duplication_null_1135 <- null_data.1135$pic[which(null_data.1135$Event=="Duplication")]
  speciation_oc_1135 <- OC_data.1135$pic[which(OC_data.1135$Event=="Speciation")]
  duplication_oc_1135 <- OC_data.1135$pic[which(OC_data.1135$Event=="Duplication")]

  ## Performing Wilcoxon tests on null and OC simulation data of trees with strong phylogenetic signals
  wilcox_2_tailed_null_1135 <- two_tailed_wilcox(speciation_null_1135,duplication_null_1135)
  wilcox_2_tailed_oc_1135 <- two_tailed_wilcox(speciation_oc_1135,duplication_oc_1135)

```

```{r Fig_S3, dpi=400,fig.width=8, fig.height=7, fig.cap= paste0("S3 Fig: Simulation analyses on 1135 trees with strong phylogenetic signals. *P* values are from Wilcoxon two-tailed tests. Dunn et al. used a cutoff of K > 0.551 to identify trees with strong phylogenetic signals. However, trees with higher K statistic can reflect trees with no phylogenetic signal when corresponding *P* values are non-significant. Considering both K statistic and *P* value, we found similar trends as was observed with 2082 trees.")}

  ## Dataframe is built based on 1135 trees with strong phylogenetic signals
  simulation_1135_new <- rbind(null_data.1135, OC_data.1135)

  ### Generating plot on simulation data for 1135 trees with strong phylogenetic signals 
  dodge <- position_dodge(width = 0.51)
  ggplot(simulation_1135_new,aes(x=label, y=pic, fill=Event)) + 
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    xlab(NULL) +
    ylab(expression(bold("Phylogenetic Independent Contrasts (PICs)"))) +
    coord_cartesian(ylim=c(0, 0.05)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9),legend.text=element_text(size=8)) +
    theme(axis.text.y= element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold"))+
    ggtitle(paste0("Plot on trees with strong phylogenetic signals"))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x = 1.04, y = 0.035, label= wilcox_2_tailed_null_1135, fontface=4) +
    annotate("text", x = 1.98, y = 0.035, label= wilcox_2_tailed_oc_1135, fontface=4)

```

```{r Fig_S4, dpi=400,fig.width=8, fig.height=7, fig.cap= paste0("S4 Fig: Re-analyses of expected variances of calibrated trees considered by Dunn et al. The expected variance plots of (A) all calibrated trees, and (B) trees with strong phylogenetic signal. The dotted line represents the mean expected variance of the events. These plots suggest that the use of duplication nodes preceding ancient speciation nodes for calibration is often erroneous and should be avoided.")}

  exp_var_tre <- paste0("A. Expected variance plot of ",length(gene_trees_calibrated), " trees", sep="")
	exp_var_trees <- paste0("B. Expected variance plot of ",length(trees_new), " trees", sep="")

	 ## Histogram plots
	 plotS4A <- ggplot2.histogram(data=check_df1, xName='exp_var',
    groupName='Event',legendPosition="right",
    alpha=0.8, addDensity=TRUE,
    addMeanLine=TRUE, meanLineColor=c("#F8766D","#00BFC4"), meanLineSize=1) +
    xlim(0, 1800)+
    theme_classic() +
    ggtitle(paste0(exp_var_tre)) +
    theme(plot.title = element_text(face = "bold"))+
    xlab("Expected variance") + 
    ylab("Density") +
    theme(legend.title=element_blank(),legend.position=c(0.7,0.8),legend.text=element_text(size=8)) +
    theme(plot.title = element_text (face="bold", 
                              colour="black", lineheight=1.0),
          axis.title.x=element_text( face="bold",
                              colour="black", hjust=0.5),
          axis.title.y=element_text(face="bold",
                              colour="black", vjust=0.5, angle=90))

  plotS4B <- ggplot2.histogram(data=check_df, xName='exp_var',
    groupName='Event',legendPosition="right",
    alpha=0.8, addDensity=TRUE,
    addMeanLine=TRUE, meanLineColor=c("#F8766D","#00BFC4"), meanLineSize=1) +
    xlim(0, 1800)+
    theme_classic() +
    ggtitle(paste0(exp_var_trees)) +
    theme(plot.title = element_text(face = "bold"))+
    xlab("Expected variance") + 
    ylab("Density") +
    theme(legend.title=element_blank(),legend.position=c(0.7,0.8),legend.text=element_text(size=8)) +
    theme(legend.position="none")+
    theme(plot.title = element_text (face="bold", 
                              colour="black", lineheight=1.0),
          axis.title.x=element_text( face="bold",
                              colour="black", hjust=0.5),
          axis.title.y=element_text(face="bold",
                              colour="black", vjust=0.5, angle=90))

    grid.arrange(plotS4A,plotS4B,nrow = 2)

```

```{r Fig_S5, dpi=400,fig.width=9, fig.height=8, fig.cap="S5 Fig: *P* value distribution plots after 100 independent runs on each set of trees. Wilcoxon two-tailed test with 95% confidence interval was used to compare the speciation and duplication contrasts after randomization tests. (A) and (B) applied to trees with at least one speciation and one duplication event. (C) and (D) applied to trees with strong phylogenetic signal. (A) and (C) randomization of trait ($\\tau$) over the trees. (B) and (D) randomization of internal node events. The inset plots show *P* values adjusted with Benjamini-Hochberg [@RN833;@RN835]. Supporting our observations of Figs 2A and 2C, all the plots confirm that the empirical result of Dunn et al. [@RN760] is not different from randomized test results."}

  fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, digit=1, scientific = TRUE)
  }

  ## Dataframe of P values for 5479 trees with at least one speciation and one duplication events 
  pval_dataframe_5479 <- data.frame(pval_tau=NA,pval_event=NA,tree_set=NA)
  pval_dataframe_5479 <- rbind(pval_dataframe_5479, data.frame(pval_tau=pval_all,pval_event=pval_all_events,tree_set="Filtered_tree"))
  pval_dataframe_5479 <- pval_dataframe_5479[-1,]
  pval_dataframe_5479$adj.ptau <- p.adjust(pval_dataframe_5479$pval_tau,method="BY")
  pval_dataframe_5479$adj.pevent <- p.adjust(pval_dataframe_5479$pval_event,method="BY")

  ## Dataframe of P values for 2082 trees with strong phylogenetic signals
  pval_dataframe_2082 <- data.frame(pval_tau=NA,pval_event=NA,tree_set=NA)
  pval_dataframe_2082 <- rbind(pval_dataframe_2082, data.frame(pval_tau=pval_2082,pval_event=pval_2082_events,tree_set="Phylosig_tree"))
  pval_dataframe_2082 <- pval_dataframe_2082[-1,]
  pval_dataframe_2082$adj.ptau <- p.adjust(pval_dataframe_2082$pval_tau,method="BY")
  pval_dataframe_2082$adj.pevent <- p.adjust(pval_dataframe_2082$pval_event,method="BY")

  ## Main plots on P value distribution after 1000 runs
  plotS5AM <- histo_plot(pval_dataframe_5479,'pval_tau','tree_set')
  plotS5AM <- plotS5AM +
           #xlim(-0.01,0.01) +
           ylim(0,100) +
           ggtitle(expression(bold(paste("A. Randomized trait, ", tau," using 5479 trees"))))+
           theme(plot.title = element_text(face="bold"))+
           theme(axis.text=element_text(size=10,face="bold"))+
           scale_x_continuous(limits = c(-0.001,0.05), labels = fancy_scientific)

  plotS5BM <- histo_plot(pval_dataframe_5479,'pval_event','tree_set')
  plotS5BM <- plotS5BM + 
           ylim(0,100) +
           ylab(" ")+
           theme (axis.title.y = element_blank())+
           ggtitle(expression(bold(paste("B. Randomized events using 5479 trees"))))+
           theme(plot.title = element_text(face="bold"))+
           theme(axis.text=element_text(size=10,face="bold"))+
           scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)

  plotS5CM <- histo_plot(pval_dataframe_2082,'pval_tau','tree_set')
  plotS5CM <- plotS5CM + 
          ylim(0,100) +
          ggtitle(expression(bold(paste("C. Randomized trait, ", tau," using 2082 trees"))))+
          theme(plot.title = element_text(face="bold"))+
          theme(axis.text=element_text(size=10,face="bold")) +
          scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)

  plotS5DM <- histo_plot(pval_dataframe_2082,'pval_event','tree_set')
  plotS5DM <- plotS5DM +
           ylim(0,100) +
           ylab(" ")+
           theme (axis.title.y = element_blank())+
           ggtitle(expression(bold(paste("D. Randomized events using 2082 trees"))))+
           theme(plot.title = element_text(face="bold"))+
           theme(axis.text=element_text(size=10,face="bold"))+
           scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)

  ##Inset plots of P value distrubution after correction for multiple testing
  plotS5AI <- histo_plot(pval_dataframe_5479,'adj.ptau','tree_set')+
              #xlim(-0.1,0.1) +
              ylim(0,100) +
              scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
              xlab(expression(bold(paste("Adjusted ", bolditalic(P)," value"))))
  plotS5BI <- histo_plot(pval_dataframe_5479,'adj.pevent','tree_set')+
              ylim(0,100) +
              scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
              xlab(expression(bold(paste("Adjusted ", bolditalic(P)," value"))))
  plotS5CI <- histo_plot(pval_dataframe_2082,'adj.ptau','tree_set')+
              ylim(0,100) +
              scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
              xlab(expression(bold(paste("Adjusted ", bolditalic(P)," value"))))
  plotS5DI <- histo_plot(pval_dataframe_2082,'adj.pevent','tree_set')+
              ylim(0,100) +
              scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
              xlab(expression(bold(paste("Adjusted ", bolditalic(P)," value"))))

  ##Drawing subplots
  plotS5A <- ggdraw() +
    draw_plot(plotS5AM,scale=0.9) +
    draw_plot(plotS5AI, x = .4, y = .4, width = .5, height = .4,scale = 1.2)

  plotS5B <- ggdraw() +
    draw_plot(plotS5BM,scale=0.9) +
    draw_plot(plotS5BI, x = .4, y = .4, width = .5, height = .4, scale=1.2)

  plotS5C <- ggdraw() +
    draw_plot(plotS5CM,scale=0.9) +
    draw_plot(plotS5CI, x = .4, y = .4, width = .5, height = .4,scale=1.2)

  plotS5D <- ggdraw() +
    draw_plot(plotS5DM,scale=0.9) +
    draw_plot(plotS5DI, x = .4, y = .4, width = .5, height = .4,scale=1.2)

  ##Arranging and placing the plots
  grid.arrange(plotS5A,plotS5B,plotS5C,plotS5D,nrow = 2, ncol=2)

```

```{r Fig_S6, dpi=400, fig.width=8, fig.height=8, fig.cap= paste("S6 Fig: Pipeline to generate different subsets of contrasts standardized BM trees of $\\tau$. The symbol '-' indicates removal of trees with the corresponding mentioned criteria. Diagnostic plot tests (discussed in details in Methods) refer to the tests recommended in prior studies [@RN781;@RN782;@RN779;@RN780;@RN783;@RN816] to identify trees where trait (in this case $\\tau$) evolution follows BM. Old duplication nodes refers to duplication events older than 370 My. Pure speciation trees indicate gene trees with no duplication event. Removal of trees without old speciation means that we keep only trees where all duplication events are more recent than the oldest speciation event, i.e. trees with age $\\le$ 296 My.")}

 ## Loading previously generated plot
 img <- png::readPNG("/Users/admin/Desktop/nwork/simulation/flowchart_S4_new.png")
 grid::grid.raster(img)

```

```{r Additional_analyses_on_2545_BM_trees, eval=T, echo=F, message=F, warning=F}

  ## Collecting tree data for all calibrated trees in a single tibble
  tree_data_2545 <- bind_rows(lapply(standarized_pic_tree_with_dup_final,tree_data_collection))
  tree_data_2545_high_dup_prop <- tree_data_2545[which(tree_data_2545$dup_prop > tree_data_2545$spe_prop),]

```

```{r Fig_S7, dpi=400,fig.width=8, fig.height=8, fig.cap="S7 Fig: Randomization test results on contrasts standardized 2545 BM trees for $\\tau$. Wilcoxon tests were used to generate *P* value distribution plots by 1000 independent runs with a node age limit of 296 My. (A) For randomized $\\tau$, Wilcoxon two-tailed tests unexpectedly showed all the values were significant. (B) *P* value of 1 in all cases of Wilcoxon one-tailed test (as performed in Dunn et al.) on randomized $\\tau$. This may be due to the fact that there is an excess proportions of speciation and NA events compared to the duplication events in these trees. By permuting $\\tau$, higher $\\tau$ values of duplication events are more likely to be assigned to the speciation and NA events. Hence, this produced higher speciation contrasts, although we controlled the bias in expected variance of older duplication events by node age limits. (C) and (D) Wilcoxon two-tailed and one-tailed tests were performed by randomization of events. After randomization, the smaller proportions of duplication events were more likely to be replaced by speciation and NA events, showing greater shifts towards excess speciation contrasts. These plots show that randomization results are different from the empirical results of Fig 3, as expected."}

  ## dataframe of P value for 2545 BM trees
  pval_2545 <- data.frame(pval_296=NA,pval_296_1sided=NA,pval_296_event=NA,pval_296_event_1sided=NA,tree_set=NA)
  pval_2545 <- rbind(pval_2545, data.frame(pval_296=pval_2545_296_2sided,pval_296_1sided=pval_2545_296_1sided,pval_296_event= pval_2545_296_event_2sided,pval_296_event_1sided=pval_2545_296_event_1sided,tree_set="2545_tree"))
  pval_2545 <- pval_2545[-1,]

  ## Plotting the randomization data result for 2545 trees 
  ## P value distribution plots of randomized Tau
  plotS7AM <- histo_plot(pval_2545,'pval_296','tree_set')
  plotS7AM <- plotS7AM + 
            ggtitle(expression(bold(paste("A. Plot on randomized ", tau))))+
            theme(plot.title = element_text(face="bold"))+    
            theme(axis.text=element_text(size=10,face="bold"))+
            scale_x_continuous(limits = c(-0.01,0.05),labels = fancy_scientific)+
            xlab(expression(bold(paste(bolditalic(P)," value (two-tailed) with age <= 296 My")))) 
  plotS7BM <- histo_plot(pval_2545,'pval_296_1sided','tree_set')+
             # xlim(0.8,1.01) +
              ggtitle(expression(bold(paste("B. Plot on randomized ", tau))))+
              theme(plot.title = element_text(face="bold"))+     
              theme(axis.text=element_text(size=10,face="bold"))+
              scale_x_continuous(limits = c(0.05,1.001),labels = fancy_scientific)+
              xlab(expression(bold(paste(bolditalic(P)," value (one-tailed) with age <= 296 My"))))

  ## P value distribution plots of randomized events
  plotS7CM <- histo_plot(pval_2545,'pval_296_event','tree_set')
  plotS7CM <- plotS7CM +
            ylim(0,150) +
            ggtitle(expression(bold(paste("C. Plot on randomized events"))))+
            theme(plot.title = element_text(face="bold"))+     
            theme(axis.text=element_text(size=10,face="bold"))+
            scale_x_continuous(labels = fancy_scientific)+
            xlab(expression(bold(paste(bolditalic(P)," value (two-tailed) with age <= 296 My"))))

  plotS7DM <- histo_plot(pval_2545,'pval_296_event_1sided','tree_set')+
              ylim(0,150) +
              ggtitle(expression(bold(paste("D. Plot on randomized events"))))+
              theme(plot.title = element_text(face="bold"))+     
              theme(axis.text=element_text(size=10,face="bold"))+
              scale_x_continuous(labels = fancy_scientific)+
              xlab(expression(bold(paste(bolditalic(P)," value (one-tailed) with age <= 296 My"))))

 gridExtra::grid.arrange(plotS7AM,plotS7BM, plotS7CM,plotS7DM,nrow = 2,ncol=2)

```

```{r Fig_S8, dpi=400,fig.width=8, fig.height=5, fig.cap=paste0("S8 Fig: Result on ", length(standarized_pic_tree_with_dup_final_new), " BM trees using maximum age of oldest duplication event as 370 My. Wilcoxon two-tailed test was used to estimate the significance level at 95% confidence interval. With different node age limits (as used in Fig 3), we found support for the ortholog conjecture. The difference between S8 Fig and Fig 3 is that in Fig 3we used the PICs of all the old duplication nodes to identify BM trees, and then we used different node age limits to test the ortholog conjecture, whereas in S8 Fig we removed the old duplication nodes (duplication age > 370 My) before identifying BM trees, and tested for the ortholog conjecture with different node age limits.")}

  ## Summerising the data slots of 2682 BM trees 
  summary_2682_BM_tree <- bind_rows(lapply(standarized_pic_tree_with_dup_final_new,summary_function))
  Contrast_brownian_2682_trees <- summary_2682_BM_tree[which(!is.na(summary_2682_BM_tree$pic)),]
  Contrast_brownian_2682_trees <- na.omit(Contrast_brownian_2682_trees[c(6,18:19,22:24)])

  ## We checked for maximum nodes contrasts for these 2682 Brownian trees, and it was less than 0.5

  ## 1st plot
  ## Using a node age limit of oldest speciation age
  Contrast_brownian_plot_296 <- Contrast_brownian_2682_trees[which(Contrast_brownian_2682_trees$node_age <= 296),]
  Contrast_brownian_plot_296$Events <- factor(Contrast_brownian_plot_296$Event, levels=c("Speciation", "Duplication"))

  # Extracting speciation and duplication PICs
  speciation_contrast_2682_trees_296 <- abs(speciation_contrast(Contrast_brownian_plot_296,'pic_Tau')) 
  duplication_contrast_2682_trees_296 <-  abs(duplication_contrast(Contrast_brownian_plot_296,'pic_Tau'))

  ## Performing test 
  wilcox_test_2682_trees_296 <- two_tailed_wilcox(duplication_contrast_2682_trees_296,speciation_contrast_2682_trees_296)

  ## 2nd plot
  ## Using a node age limit of considering duplication age is 25% higher than oldest speciation age
  Contrast_brownian_plot_370 <- Contrast_brownian_2682_trees[which(Contrast_brownian_2682_trees$node_age <= 370),]
  Contrast_brownian_plot_370$Events <- factor(Contrast_brownian_plot_370$Event, levels=c("Speciation", "Duplication"))

  # Separating speciation and duplication PICs
  speciation_contrast_2682_trees_370 <- abs(speciation_contrast(Contrast_brownian_plot_370,'pic_Tau')) 
  duplication_contrast_2682_trees_370 <-  abs(duplication_contrast(Contrast_brownian_plot_370,'pic_Tau'))

  ## Performing test 
  wilcox_test_2682_trees_370 <- two_tailed_wilcox(duplication_contrast_2682_trees_370,speciation_contrast_2682_trees_370)

  ## Making a  dataframe for plotting
  plotting_data <- data.frame(pic=NA, group=NA, Event=NA) ## declaring data frame
  plotting_data <- rbind(plotting_data, data.frame(pic=abs(Contrast_brownian_plot_296$pic_Tau),  group="Age <= 296 My", Event=Contrast_brownian_plot_296$Events))
  plotting_data <- rbind(plotting_data, data.frame(pic=abs(Contrast_brownian_plot_370$pic_Tau),  group="Age <= 370 My", Event=Contrast_brownian_plot_370$Events))
  plotting_data <- plotting_data[-1,]
  plotting_data$Event <- factor(plotting_data$Event, levels=c("Speciation", "Duplication"))

  ## Boxplot 
  dodge <- position_dodge(width = 0.51)
    ggplot(plotting_data, aes(group, pic, fill = Event)) +
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
   #geom_jitter(aes(colour = Event),position=dodge, alpha=0.02) +
    xlab(NULL) +
    ylab(expression(bold(paste("PIC of ",tau)))) +
    coord_cartesian(ylim=c(0, 0.06)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9)) +
    #theme(legend.position="none")+
    theme(axis.text.y = element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold")) +
    #ggtitle(paste0(standardized_trees))+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x=1.04, y=0.05, label=wilcox_test_2682_trees_296, fontface=4)+ 
    annotate("text", x=1.98, y=0.05, label=wilcox_test_2682_trees_370, fontface=4)

```

```{r Fig_S9, dpi=400,fig.width=8, fig.height=5, fig.cap=paste0("S9 Fig: Distribution of proportions of events in 806 BM trees. These trees have a median of ", median_number_of_events, " internal nodes. Most of the trees have speciation nodes whose descendants have the same clade name. Hence, Dunn et al. [@RN760] changed such nodes from speciation events to NA to avoid problem with tree calibration. From the figure, it is clear that the proportion of speciation events is much higher than that of the duplication events. For these trees, the proportions of NA events (median:  ", median_proporion_of_NA_event, ") are even higher than the proportion of duplication events (median: ",median_proporion_of_duplication,"). The dotted line represents the median proportions of the events.")}

  ## Creating a dataframe for plotting the proportions of events
  df_806 <- data.frame(proportion_event=NA,Event=NA)
  df_806 <- rbind(df_806, data.frame(proportion_event=tree_data_806$spe_prop,Event="Speciation"))
  df_806 <- rbind(df_806, data.frame(proportion_event=tree_data_806$dup_prop,Event="Duplication"))
  df_806 <- rbind(df_806, data.frame(proportion_event=tree_data_806$NA_prop,Event="NA"))
  df_806 <- df_806[-1,]
  df_806$Event <- factor(df_806$Event, levels=c("Speciation", "Duplication","NA"))

  ## Generating plot to show distribution of events for all calibrated trees
  ggplot2_histogram_mod(data=df_806, xName='proportion_event',
    groupName='Event',groupColors=c("#F8766D","#00BFC4","#C77CFF"), legendPosition="right",
    alpha=1, addDensity=TRUE,
    addMedianLine=TRUE, medianLineColor=c("#F8766D","#00BFC4","#C77CFF"), medianLineSize=1) +
    theme_classic() +
    ggtitle(expression(bold("Event proportion plot for 806 BM trees")))+
    theme(plot.title = element_text(face="bold"))+
    xlab("Proportion of events") + 
    ylab("Density") +
    theme(legend.title=element_blank(),legend.position=c(0.9,0.8),legend.text=element_text(size=8)) +
    theme(axis.text=element_text(size=10,face="bold")) +
    theme(plot.title = element_text (face="bold", 
                              colour="black", lineheight=1.0),
          axis.title.x=element_text( face="bold",
                              colour="black", hjust=0.5),
          axis.title.y=element_text(face="bold",
                              colour="black", vjust=0.5, angle=90))

```

```{r Fig_S10, eval = T, dpi=400,fig.width=10, fig.height= 9, fig.cap= paste("S10 Fig:  Re-analyses of 3151 trees called as following a BM model by Dunn et al. (A) Wilcoxon two-tailed test was used to obtain the significance level using empirical data. The box plot shows that there is no support for the ortholog conjecture for recent duplicates (age $\\le$ 296 My), but that there is support for the older duplicates. (B) and (C) *P* values estimated using randomized $\\tau$ (B), and using randomized events (C), with a node age limit of 296 My to test if there was any difference in contrasts between speciation and duplication events, and plotted based on 100 independent *P* values. The main plots in (B) and (C) demonstrate that speciation and duplication contrasts differ significantly in each randomization test. The inset plots of (B) and (C) with Wilcoxon one-tailed test to check if duplication contrasts was higher than the speciation contrasts, confirm the opposite trend, as was observed with empirical data with node age $\\le$ 296 My. This suggests that contrasts are not adequately standardized for these trees as per BM, and that the inference based on these trees can be misleading.")}

  ## Making test on Dunn identified BM trees with node age limits
  speciation_BM_pic <- nodes_contrast_filtered_bm$pic[which(nodes_contrast_filtered_bm$Event=="Speciation")]
  duplication_ancient_BM_pic <- nodes_contrast_filtered_bm$pic[which(nodes_contrast_filtered_bm$Event=="Duplication" & nodes_contrast_filtered_bm$node_age > 296 )]
  duplication_recent_BM_pic <- nodes_contrast_filtered_bm$pic[which(nodes_contrast_filtered_bm$Event=="Duplication" & nodes_contrast_filtered_bm$node_age <= 296)]

  ## Performing test 
  wilcox_test_BM_ancient_dup <- two_tailed_wilcox(duplication_ancient_BM_pic,speciation_BM_pic)
  wilcox_test_BM_recent_dup <- two_tailed_wilcox(duplication_recent_BM_pic ,speciation_BM_pic)

  ## Dataframe of speciation and ancient duplication events
  BM_ancient <- data.frame(pic_tau= NA,Event=NA)
  BM_ancient <- rbind(BM_ancient, data.frame(pic_tau= speciation_BM_pic,Event="Speciation"))
  BM_ancient <- rbind(BM_ancient, data.frame(pic_tau= duplication_ancient_BM_pic,Event="Duplication"))
  BM_ancient <- BM_ancient[-1,]
  BM_ancient$Event <- factor(BM_ancient$Event, levels=c("Speciation", "Duplication"))

  ## Creating a dataframe of speciation and recent duplication
  BM_recent <- data.frame(pic_tau= NA,Event=NA)
  BM_recent <- rbind(BM_recent, data.frame(pic_tau= speciation_BM_pic, Event="Speciation"))
  BM_recent <- rbind(BM_recent, data.frame(pic_tau= duplication_recent_BM_pic,Event="Duplication"))
  BM_recent <- BM_recent[-1,]
  BM_recent$Event <- factor(BM_recent$Event, levels=c("Speciation", "Duplication"))

  ## Creating a dataframe for plotting
  plot_BM <- data.frame(pic=NA, group=NA, Event=NA) ## declaring data frame
  plot_BM <- rbind(plot_BM, data.frame(pic=BM_recent$pic_tau, group="Duplicate age <= 296 My", Event=BM_recent$Event))
  plot_BM <- rbind(plot_BM, data.frame(pic=abs(BM_ancient$pic_tau), group="Duplicate age > 296 My", Event=BM_ancient$Event))
  plot_BM <- plot_BM[-1,]
  plot_BM$Event <- factor( plot_BM$Event, levels=c("Speciation", "Duplication"))

  ## Plotting the data
  dodge <- position_dodge(width = 0.51)

  plotS10A <- ggplot(plot_BM, aes(group, pic, fill = Event)) +
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
   #geom_jitter(aes(colour = Event),position=dodge, alpha=0.02) +
    xlab(NULL) +
    ylab(expression(bold(paste("PIC of ",tau)))) +
    coord_cartesian(ylim=c(0, 0.06)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9)) +
    #theme(legend.position="none")+
    theme(axis.text.y = element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold")) +
    ggtitle("A. Analyses on BM trees")+
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x=1.04, y=0.05, label=wilcox_test_BM_recent_dup, fontface=4) +
    annotate("text", x=1.98, y=0.05, label=wilcox_test_BM_ancient_dup, fontface=4)

   ## Analyses on Pvalue on trees called BM by Dunn et al.
  pval_BM_dunn <- data.frame(pval_296_2sided=NA,pval_296_1sided=NA,pval_296_2sided_event=NA,pval_296_1sided_event=NA,tree_set=NA)
  pval_BM_dunn <- rbind(pval_BM_dunn, data.frame(pval_296_2sided=pval_BM_Dunn_296_2sided,pval_296_1sided=pval_BM_Dunn_296_1sided,pval_296_2sided_event= pval_BM_Dunn_296_2sided_event,pval_296_1sided_event=pval_BM_Dunn_296_1sided_event,tree_set="Dunn_BM_trees"))
  pval_BM_dunn <- pval_BM_dunn[-1,]

  ## Plotting the randomization data result for BM trees  of Dunn et al.
  ## Main plots on P value distribution after 1000 runs
  plotS10BM <- histo_plot(pval_BM_dunn,'pval_296_2sided','tree_set')
  plotS10BM <- plotS10BM +
            ylim(0,100) +
            scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
            ggtitle(expression(bold(paste("B. Plot on BM trees using randomized ", tau))))+
            theme(plot.title = element_text(face="bold"))+    
            theme(axis.text=element_text(size=9,face="bold"))+
            xlab(expression(bold(paste(bolditalic(P)," value (two-tailed) with age <= 296 My")))) 
  plotS10CM <- histo_plot(pval_BM_dunn,'pval_296_2sided_event','tree_set')
  plotS10CM <- plotS10CM +
            ylim(0,100) +
            scale_x_continuous(limits = c(-0.001,0.05),labels = fancy_scientific)+
            ggtitle(expression(bold(paste("C. Plot on BM trees using randomized events"))))+
            theme(plot.title = element_text(face="bold"))+     
            theme(axis.text=element_text(size=10,face="bold"))+
            xlab(expression(bold(paste(bolditalic(P)," value (two-tailed) with age <= 296 My"))))

  ## Inset plot 
  plotS10IB <- histo_plot(pval_BM_dunn,'pval_296_1sided','tree_set')+
              #xlim(0.9,1.01) +
              scale_x_continuous(limits = c(0.05,1.001),labels = fancy_scientific)+
              ylim(0,100) +
              xlab(expression(bold(paste(bolditalic(P)," value (one-tailed) with age <= 296 My")))) 
  plotS10IC <- histo_plot(pval_BM_dunn,'pval_296_1sided_event','tree_set')+
              scale_x_continuous(limits = c(0.05,1.001),labels = fancy_scientific)+
              ylim(0,100) +
              xlab(expression(bold(paste(bolditalic(P)," value (one-tailed) with age <= 296 My"))))

  ## Preparing final plots and arranging them
  plotS10B <- ggdraw() +
    draw_plot(plotS10BM) +
    draw_plot(plotS10IB, x = .45, y = .4, width = .5, height = .57, scale=0.8)

  plotS10C <- ggdraw() +
    draw_plot(plotS10CM) +
    draw_plot(plotS10IC, x = .45, y = .4, width = .5, height = .57,scale=0.8)

 grid.arrange(plotS10A,plotS10B,plotS10C,nrow = 3)

```

```{r Fig_S11, dpi=400,fig.width=8, fig.height=5, fig.cap= paste0("S11 Fig: The ortholog conjecture test on Dunn et al. identified BM trees, after diagnostic tests. Among 3151 trees, ", length(BM_tree_diagnostic_test_final), " trees passed the diagnostic tests of BM. Wilcoxon two-tailed test was used to estimate the significance level at 95% confidence interval. We did not exclude pure speciation trees before the contrasts analyses. With node age limit (age $\\le$ 296 My) as well as without, we found support for the ortholog conjecture.")}

  ## summerising the results for newly identified BM trees 
  summary_new_BM_tree <- bind_rows(lapply(BM_tree_diagnostic_test_final,summary_function))
  Contrast_brownian_new_trees <- summary_new_BM_tree[which(!is.na(summary_new_BM_tree$pic)),]
  Contrast_brownian_new_trees <- na.omit(Contrast_brownian_new_trees[c(6,18:19,22:24)])
  Contrast_brownian_new_trees$Events <- factor(Contrast_brownian_new_trees$Event, levels=c("Speciation", "Duplication"))
  ## We checked for maximum nodes contrasts for these 2412 Brownian trees, and it was less than 0.5

  ## We use a node age limit of oldest speciation age
  BM_contrast_296 <- Contrast_brownian_new_trees[which(Contrast_brownian_new_trees$node_age <= 296),]
  BM_contrast_296$Events <- factor(BM_contrast_296$Event, levels=c("Speciation", "Duplication"))

  # Separating speciation and duplication contrast data with an age limit of 296 My
  speciation_contrast_new_296 <- speciation_contrast(BM_contrast_296,'pic') 
  duplication_contrast_new_296 <- duplication_contrast(BM_contrast_296,'pic')

  # Separating speciation and duplication contrast data with no age limit
  speciation_contrast_new_all <- abs(speciation_contrast(Contrast_brownian_new_trees,'pic')) 
  duplication_contrast_new_all <- abs(duplication_contrast(Contrast_brownian_new_trees,'pic')) 
   ## Performing test 
   wilcox_test_new_trees_296 <- two_tailed_wilcox(duplication_contrast_new_296,speciation_contrast_new_296)
   wilcox_test_new_trees_all <- two_tailed_wilcox(duplication_contrast_new_all,speciation_contrast_new_all)

  ## Making a  dataframe for plotting
  data_plot <- data.frame(pic=NA, group=NA, Event=NA) ## declaring data frame
  data_plot <- rbind(data_plot, data.frame(pic=abs(BM_contrast_296$pic),  group="Age <= 296 My", Event=BM_contrast_296$Events))
  data_plot <- rbind(data_plot, data.frame(pic=abs(Contrast_brownian_new_trees$pic),  group="No age limit", Event=Contrast_brownian_new_trees$Events))
  data_plot <- data_plot[-1,]
  data_plot$Event <- factor(data_plot$Event, levels=c("Speciation", "Duplication"))

  ## Plotting the data
  dodge <- position_dodge(width = 0.51)

    ggplot(data_plot, aes(group, pic, fill = Event)) +
    geom_boxplot(width=0.5,outlier.colour=NA, position=dodge, notch=T) +
    xlab(NULL) +
    ylab(expression(bold(paste("PIC of ",tau)))) +
    coord_cartesian(ylim=c(0, 0.06)) +
    theme_classic()+
    theme(legend.title=element_blank(),legend.position=c(0.9,0.9)) +
    theme(axis.text.y = element_text(size=10)) +
    theme(axis.text.x = element_text(size=10,face="bold")) +
    theme(plot.title = element_text(face = "bold"))+
    annotate("text", x=1.04, y=0.05, label=wilcox_test_new_trees_296, fontface=4)+ 
    annotate("text", x=1.98, y=0.05, label=wilcox_test_new_trees_all, fontface=4)

```

\vspace{0.5cm}

\pagebreak

### Reference